------------------------------------------------------------------------

---
title: "gene_module"
output: pdf_document
---

```{r,message=FALSE, warning=FALSE, results='hide' }
setwd("/project/xuanyao/jiaming/Getting_started")

library(presto)
library(sceptre)
library(sceptredata)
library(readr)
library(dplyr)
library(Matrix)
library(parallel)
library(BH)
library(Rcpp)
library(peakRAM)
library(sgt)

sourceCpp("code/fit_skew_normal.cpp")

```

```{r, message=FALSE, warning=FALSE, results='hide'}
Morris_data <- T

setwd("/project/xuanyao/jiaming/Getting_started")


if (Morris_data) {
  directories <- "data/STINGseq-v1_GDO"
  grna_target_data_frame <- data.frame(read_csv(paste0(directories,"/grna_target_data_frame.csv")))
  sceptre_object <- import_data_from_cellranger(
  directories = directories,
  moi = "high",
  grna_target_data_frame = grna_target_data_frame
)
} else {
  directories <- paste0(system.file("extdata", package = "sceptredata"), "/highmoi_example/gem_group_", 1:2)
  data(grna_target_data_frame_highmoi)
  sceptre_object <- import_data_from_cellranger(
  directories = directories,
  moi = "high",
  grna_target_data_frame = grna_target_data_frame_highmoi
)
}


positive_control_pairs <- construct_positive_control_pairs(sceptre_object)

discovery_pairs_cis <- construct_cis_pairs(
  sceptre_object = sceptre_object,
  positive_control_pairs = positive_control_pairs,
  response_position_data_frame=gene_position_data_frame_grch37,
  distance_threshold = 5e5
)

discovery_pairs_trans <- construct_trans_pairs(
  sceptre_object = sceptre_object,
  positive_control_pairs = positive_control_pairs,
  pairs_to_exclude = "pairs_containing_pc_targets"
)
sceptre_object <- sceptre_object |> # |> is R's base pipe, similar to %>%
  set_analysis_parameters(
    discovery_pairs=discovery_pairs_trans, 
    positive_control_pairs=positive_control_pairs, 
    side="both") |>
  assign_grnas()
```

## Modifying and overwrite the functions in Presto library to do wilcoxon test

```{r}
compute_ustat <- function(Xr, cols_list, n1n2_list, group_size_list) {
  total_loops <- length(cols_list) # Total number of loops
  loop_counter <- 1 # Initialize loop counter
  
  mapply(function(cols, n1n2, group_size) {
    cat("Compute wilcox test statisitcs for gRNA #", loop_counter, "of", total_loops,".    ", round(loop_counter / total_loops * 100, 2)
,"% of total progress" , "\n") 
    loop_counter <<- loop_counter + 1 # Increment the counter
    
    grs <- sumGroups(Xr, cols)
    
    if (is(Xr, "dgCMatrix")) {
      gnz <- (group_size - nnzeroGroups(Xr, cols))
      zero_ranks <- (nrow(Xr) - diff(Xr@p) + 1) / 2
      ustat <- t((t(gnz) * zero_ranks)) + grs - group_size * (group_size + 1) / 2
    } else {
      ustat <- grs - group_size * (group_size + 1) / 2
    }
    
    return(ustat)
  }, cols_list, n1n2_list, group_size_list, SIMPLIFY = FALSE)
}



compute_pval <- function(ustat_list, ties, N, n1n2_list) {
  total_loops <- length(ustat_list) # Total number of loops
  loop_counter <- 1 # Initialize loop counter
  mapply(function(ustat, n1n2) {
    cat("Compute wilcox p value for gRNA #", loop_counter, "of", total_loops,".    ", round(loop_counter / total_loops * 100, 2)
,"% of total progress" , "\n") 
    loop_counter <<- loop_counter + 1 # Increment the counter
    z <- ustat - 0.5 * n1n2
    z <- z - sign(z) * 0.5
    .x1 <- N ^ 3 - N
    .x2 <- 1 / (12 * (N^2 - N))
    
    rhs <- lapply(ties, function(tvals) {
      (.x1 - sum(tvals ^ 3 - tvals)) * .x2
    }) %>% unlist
    
    usigma <- sqrt(matrix(n1n2, ncol = 1) %*% matrix(rhs, nrow = 1))
    z <- t(z / usigma)
    
    pvals <- matrix(2 * pnorm(-abs(as.numeric(z))), ncol = ncol(z))
    return(pvals)
  }, ustat_list, n1n2_list, SIMPLIFY = FALSE)
}



tidy_results_reduced <- function(wide_res, features, groups) {
  res <- Reduce(cbind, lapply(wide_res, as.numeric)) %>% data.frame()
  colnames(res) <- names(wide_res)
  
  # Create the 'feature' and 'group' columns
  res$feature <- rep(features, times = length(groups))
  res$group <- rep(groups, each = length(features))
  
  # Filter only the first group
  first_group <- groups[1]
  res <- res %>% dplyr::filter(group == first_group)
  
  # Select the relevant columns
  res %>% dplyr::select(
    feature,
    statistic,
    pval
  )
}


wilcoxauc <- function(X, ...) {
    UseMethod("wilcoxauc")
}

wilcoxauc.default <- function(X, y, groups_use = NULL, verbose = TRUE, ...) {
    ## Check and possibly correct input values
    if (is(X, "dgeMatrix")) X <- as.matrix(X)
    if (is(X, "data.frame")) X <- as.matrix(X)
    if (is(X, "dgTMatrix")) X <- as(X, "dgCMatrix")
    if (is(X, "TsparseMatrix")) X <- as(X, "dgCMatrix")
    if (ncol(X) != length(y[[1]])) stop("number of columns of X does not
                                match length of y")
    if (!is.null(groups_use)) {
        idx_use <- which(y %in% intersect(groups_use, y))
        y <- y[idx_use]
        X <- X[, idx_use]
    }

    y  <- lapply(y, factor)
  

    group.size <- lapply(y, function(vec) as.numeric(table(vec)))
  

    if (is.null(row.names(X))) {
        row.names(X) <- paste0("Feature", seq_len(nrow(X)))
    }

    ## Compute primary statistics
    
    n1n2 <- lapply(group.size, function(gs) gs * (ncol(X) - gs))
    
    if (is(X, "dgCMatrix")) {
        rank_res <- rank_matrix(Matrix::t(X))
    } else {
        rank_res <- t(rank_matrix(t(X)))
    }
      
    ustat <- compute_ustat(rank_res$X_ranked, y, n1n2, group.size)
    pvals <- compute_pval(ustat, rank_res$ties, ncol(X), n1n2)
  

    
    results_list <- lapply(seq_along(ustat), function(i) {
  # Create a res_list for each i-th vector in ustat and pvals
    res_list <- list(
    pval = pvals[[i]],
    statistic = t(ustat[[i]])
  )
  
    tidy_results_reduced(res_list, row.names(X), levels(y[[i]]))
})

# results_list is now a list of the tidy_results_reduced outputs
return(results_list)
}
```

## Estimate Propensity using logistic regression

```{r}
estimate_propensity <- function (grna_assignment_matrix, covariate) {
  
  estimated_prop <- matrix(NA, nrow = nrow(grna_assignment_matrix), ncol = ncol(grna_assignment_matrix))

# Loop through each row of the grna_assignment_matrix
for (i in 1:nrow(grna_assignment_matrix)) {
  
  # Extract the current row from the grna_assignment_matrix
  grna_row <- as.numeric(grna_assignment_matrix[i, ])
  
  # Fit logistic regression model using covariates and the current row
  glm_fit <- glm(grna_row ~ ., family = binomial(link = "logit"), data = covariate)
  
  # Predict the fitted means (estimated probabilities) for this row
  estimated_prop[i, ] <- predict(glm_fit, type = "response")
}
  return(estimated_prop)
}
```

## Permutation/Conditional resampling on gRNA assignment and combine the original matrix and permuted matrix as a big matrix

```{r}
resample_and_combine<- function(grna_assignment_matrix, n_permute) {
  propensity_matrix <- estimate_propensity(grna_assignment_matrix , covariate)
  
  propensity <- do.call(rbind, replicate(n_permute, propensity_matrix, simplify = FALSE))
  
  
    resampled_mat <- matrix(rbinom(n = length(propensity), 
                                  size = 1, 
                                  prob = as.vector(propensity)),
                           nrow = nrow(propensity), 
                           ncol = ncol(propensity))
  
    rownames(resampled_mat) <- rep(rownames(grna_assignment_matrix),n_permute)
    combined_mat <- rbind(grna_assignment_matrix, resampled_mat)
    
  return(combined_mat)
}

permute_and_combine <- function(grna_assignment_matrix, n_permute) {
  
  permute_list <- list()
  
  for (i in 1:n_permute) {
    # Generate a permutation of the column indices
    permuted_indices <- sample(ncol(grna_assignment_matrix))
    
    # Apply the permutation to the columns
    permute_list[[i]] <- grna_assignment_matrix[, permuted_indices]
  
  }
  combined_mat <-do.call(rbind, permute_list)
  combined_matrix_with_original <- rbind(grna_assignment_matrix, combined_mat)
  return(combined_matrix_with_original)
}

permutation_or_resampling <- function(grna_assignment_matrix, n_permute, use_resample=F){
  if(use_resample){
    return(resample_and_combine(grna_assignment_matrix, n_permute))
  } else{
    return(permute_and_combine(grna_assignment_matrix, n_permute))
  }
}
```

## Revised functions to calculate both original statistics and permuted statistics at a time

```{r}


slice_to_list <- function(large_matrix, n_permute) {
  # Initialize an empty list to store the smaller matrices
  sliced_matrices <- list()
  
  # Calculate the number of rows for each smaller matrix
  slice_size <- n_rows

 for (i in 1:(n_permute)) {
    start_row <- (i - 1) * slice_size + 1
    end_row <- i * slice_size
    
    sliced_matrices[[i]] <- large_matrix[start_row:end_row, ]
  }
  
  return(sliced_matrices)
}
t_test <- function(response_matrix, grna_assignment_matrix) {

  response_matrix <- as.matrix(response_matrix)
  grna_assignment_matrix <- as.matrix(grna_assignment_matrix)

  n_cells <- ncol(response_matrix)
  n_genes <- nrow(response_matrix)
  all_cells <- seq_len(n_cells)
  n_rows <- nrow(grna_assignment_matrix)

  response_matrix_t <- t(response_matrix)
  response_matrix_t_squared <-response_matrix_t^2

  grna_infected_cell_count <- as.vector(grna_assignment_matrix %*% matrix(rep(1, n_cells), nrow = n_cells))

  mean_expression_trt <- grna_assignment_matrix %*% response_matrix_t / grna_infected_cell_count
  mean_expression_squared_trt <- grna_assignment_matrix %*% response_matrix_t_squared / grna_infected_cell_count
  var_expression_trt <- mean_expression_squared_trt - mean_expression_trt^2

  grna_control_cell_count <- n_cells - grna_infected_cell_count
  mean_expression_control <- (1 - grna_assignment_matrix) %*% response_matrix_t / grna_control_cell_count
  mean_expression_squared_control <- (1 - grna_assignment_matrix) %*% response_matrix_t_squared / grna_control_cell_count
  var_expression_control <- mean_expression_squared_control - mean_expression_control^2

  t_statistics <- (mean_expression_trt - mean_expression_control) / sqrt(var_expression_trt / grna_infected_cell_count + var_expression_control / grna_control_cell_count)

  return(t_statistics)
}


wilcox_test <- function(response_matrix, grna_assignment_matrix){
  
  n_rows <- nrow(grna_assignment_matrix)
  
  response_dgC <- as(as.matrix(response_matrix),"dgCMatrix")
  # response_dgC <- as.matrix(response_matrix)

  grna_assignment_list <- split(grna_assignment_matrix, row(grna_assignment_matrix))
  
  wilcox_result <- wilcoxauc(response_dgC,grna_assignment_list)

  names(wilcox_result) <- rownames(grna_assignment_matrix)
  
  statistic_list <- lapply(wilcox_result, function(df) df$pval)


  statistic_matrix <- do.call(rbind, statistic_list)
  colnames(statistic_matrix) <- rownames(response_matrix)
  
  return(statistic_matrix)
}
```

## Combining test statistics within gene modules

```{r}
convert_gene_list_to_matrix <-function(response_matrix, gene_module){
  
  genes_in_response <- rownames(response_matrix)
  
  # Initialize a matrix to store the result
  result_matrix <- matrix(0, nrow = nrow(response_matrix), ncol = length(gene_module))
  
  # Loop over each gene module (i.e., each list of genes in gene_module)
  for (i in seq_along(gene_module)) {
    # Get the current gene module (a vector of genes)
    current_module <- gene_module[[i]]
    
    # For each gene in the response_matrix, check if it's in the current gene module
    result_matrix[, i] <- as.numeric(genes_in_response %in% current_module)
  }
  
  colnames(result_matrix) <- names(gene_module)
  
  # Return the resulting matrix
  return(result_matrix)
}

combine_columns_from_matrices <- function(matrices,gene_module_matrix) {
  # Combine columns from the input matrices column by column
  combined <- do.call(cbind, lapply(1:ncol(matrices[[1]]), function(i) {
    # Extract the i-th column from each matrix and combine them
    cbind(sapply(matrices, function(mat) mat[, i]))
  }))
  colnames(combined) <- rep(colnames(gene_module_matrix),each=length(matrices))
  return(combined)
}

combine_test_statistics <- function(test_result, gene_module_matrix){
    
    power_sum_1 <- test_result%*%gene_module_matrix
    power_sum_2 <- test_result^2%*%gene_module_matrix
    power_sum_3 <- test_result^3%*%gene_module_matrix
    power_sum_4 <- test_result^4%*%gene_module_matrix
    power_sum_5 <- test_result^5%*%gene_module_matrix
    power_sum_6 <- test_result^6%*%gene_module_matrix
    
    combined_matrix <- combine_columns_from_matrices(list(power_sum_1,power_sum_2,power_sum_3,power_sum_4,power_sum_5,power_sum_6), gene_module_matrix)
  
    return(combined_matrix)

}

convert_matrix_to_tensor <- function(matrix, n_permute, grna_assignment_matrix) {
  
  # Get the number of rows in grna_assignment_matrix, this will be the number of slices (depth)
  depth <- nrow(grna_assignment_matrix)
  
  # Initialize an empty tensor with dimensions (n_permute x ncol(matrix) x depth)
  tensor <- array(0, dim = c((n_permute+1), ncol(matrix), depth))
  
  # Loop over the number of slices (depth)
  for (i in 1:depth) {
    # For each slice, extract the corresponding rows for each group (NTC-1, NTC-2, ..., NTC-n_permute)
    row_indices <- seq(i, nrow(matrix), by = depth)
    
    # Assign the extracted rows to the i-th slice of the tensor
    tensor[,,i] <- matrix[row_indices, ]
  }
  
  dimnames(tensor) <- list(
    
    c("Original", paste0("Permutation", 1:n_permute)),    
    # Second dimension: Column names from the input matrix
    colnames(matrix),
    
    # Third dimension: Slice names based on grna_assignment_matrix (e.g., the row names or indices)
    rownames(grna_assignment_matrix)
  )
  return(tensor)
}
```

## Compute p-values of combined test statistics

```{r}
rank_tensor <- function(tensor) {
  # Get the dimensions of the input tensor
  dim_tensor <- dim(tensor)
  
  # Initialize an empty tensor to store the ranked values with the same dimensions
  ranked_tensor <- array(NA, dim = dim_tensor, dimnames = dimnames(tensor))
  
  # Use mclapply to parallelize the operation across the 3rd dimension (slices)
  ranked_slices <- mclapply(1:dim_tensor[3], function(i) {
    # Rank each slice (2D matrix) of the tensor
    ranked_slice <- rank_matrix(t(tensor[,,i]))$X_ranked
    print(paste("Complete ranking test statistics for grna", i))
    return(ranked_slice)
  }, mc.cores = parallel::detectCores() - 1)  # Using all available cores except one
  
  # Assign the ranked slices back to the ranked_tensor
  for (i in 1:dim_tensor[3]) {
    ranked_tensor[,,i] <- ranked_slices[[i]]
  }
  
  return(ranked_tensor)
}


compute_p_value_from_tensor <- function(tensor,n_permute,side_code){
  
  rank<- rank_tensor(tensor)
  if (side_code==1){ #right side
    rank <- (n_permute+2-rank)
  } 
  return(rank/(n_permute+1))
}

compute_p_value_from_skew_normal_2 <- function(tensor, side_code) {
  # Get dimensions and dimnames of the tensor
  dim_tensor <- dim(tensor)
  dimnames_tensor <- dimnames(tensor)
  
  # Initialize a tensor to store p-values with the same dimensions and dimnames
  p_value_tensor <- array(NA, dim = c(dim_tensor[1], dim_tensor[2], dim_tensor[3]), dimnames = dimnames_tensor)
  
  # Apply over each slice (i.e., the 3rd dimension)
  result <- apply(tensor, 3, function(slice) {
    # Parallelize over the second dimension (columns) for each slice
    mclapply(1:dim_tensor[2], function(i) {
      # Apply the function to the column vector slice[,i]
      fit_and_evaluate_skew_normal(slice[,i], side_code, F)
    }, mc.cores = parallel::detectCores() - 1)  # Adjust the number of cores used
  })
  
  # Rearrange the results back into the p_value_tensor
  for (j in 1:dim_tensor[3]) {
    for (i in 1:dim_tensor[2]) {
      p_value_tensor[, i, j] <- result[[j]][[i]]
    }
  }
  
  return(p_value_tensor)
}

compute_p_value_from_skew_normal <- function(tensor, side_code) {
  # Get dimensions and dimnames of the tensor
  dim_tensor <- dim(tensor)
  dimnames_tensor <- dimnames(tensor)
  
  # Initialize an empty tensor with the same dimensions and dimnames
  p_value_tensor <- array(NA, dim = dim_tensor, dimnames = dimnames_tensor)
  
  # Use mclapply to apply the function in parallel across all slices
  result <- mclapply(1:dim_tensor[3], function(j) {
    # Apply the function across the first dimension (i.e., columns of tensor[,i,j])
    apply(tensor[,,j], 2, function(column_vector) {
      tryCatch({
        # Apply the fit_and_evaluate_skew_normal function to each column vector
        fit_and_evaluate_skew_normal(column_vector, side_code, F)
      }, error = function(e) {
        # Catch and handle any errors, print message, return NA
        message("Error in slice ", j, " for column vector: ", e)
        return(rep(NA, length(column_vector)))  # Return NA for the column if an error occurs
      })
    })
  }, mc.cores = parallel::detectCores() - 1)  # Adjust the number of cores used
  
  # Combine the results back into the tensor
  for (j in 1:dim_tensor[3]) {
    p_value_tensor[,,j] <- result[[j]]
  }
  
  return(p_value_tensor)
}
```

```{r}
# # Create a function to calculate the tail probability for rows with the same name
# compute_p_val <- function(group) {
#   # Create an empty matrix to store p-values
#   p_values <- matrix(NA, nrow = nrow(group), ncol = ncol(group))
#   colnames(p_values) <- colnames(group)
#   
#   # Loop over each row in the group (rows with the same name)
#   for (i in 1:nrow(group)) {
#     current_row <- group[i, ] 
#     other_rows <- group[-i, ]  
#     
#      current_row_copy <- do.call(rbind, replicate(nrow(other_rows), current_row,    simplify = FALSE))
# 
# 
#     comparison <- other_rows > current_row_copy
# 
#  p_values[i, ] <- colSums(comparison) / nrow(other_rows)
#  
#   }
#   rownames(p_values) <- rownames(group)
#   
#   return(p_values)
# }
```

```{r}
# compute_p_val_one_vs_group <- function(original,permuted,n_permute, bigger ){
#   
#   copy <- do.call(rbind, replicate(n_permute, original, simplify = FALSE))
# 
# 
#  unique_row_names <- unique(rownames(copy))
# 
# # Initialize an empty list to store the comparison results
# comparison_results <- list()
# 
# # Loop over each unique row name
# for (row_name in unique_row_names) {
#   # Extract the rows with the current row name from both matrices
#   copy_rows <- copy[rownames(copy) == row_name, ]
#   permuted_rows <- permuted[rownames(permuted) == row_name,  ]
#   if(bigger){
#     comparison <- permuted_rows > copy_rows
#   } else {
#     comparison <- permuted_rows < copy_rows
#   }
#   
#   
#    p_values <- colSums(comparison) / n_permute
#   # Store the result in the list
#   comparison_results[[row_name]] <- p_values
# }
# 
# # Combine the results into a single matrix using cbind
# p_val <- do.call(rbind, comparison_results)
# }
```

```{r}
# min_p_val_2 <- function(p_val_tensor, gene_module_index_matrix) {
#   # Get the dimensions of the input tensor
#   dim_tensor <- dim(p_val_tensor)
# 
#   # Get the number of groups from the gene_modules matrix (n_groups)
#   n_groups <- ncol(gene_module_index_matrix)
# 
#   # Initialize an empty tensor to store the results with the new dimensions
#   min_p_tensor <- array(NA, dim = c(dim_tensor[1], n_groups, dim_tensor[3]), dimnames = list(dimnames(p_val_tensor)[[1]], colnames(gene_module_index_matrix), dimnames(p_val_tensor)[[3]]))
# 
#   # Loop over each slice of the tensor (each slice corresponds to a 2D matrix)
#   for (i in 1:dim_tensor[3]) {
#     # Extract the current slice (2D matrix)
#     current_slice <- p_val_tensor[,,i]
# 
#     # Extract the column names of current_slice
#     slice_colnames <- colnames(current_slice)
# 
#     # Loop over each group/module in gene_modules
#     for (j in 1:n_groups) {
#       # Extract the column names for the current group from gene_modules
#       group_colnames <- colnames(gene_module_index_matrix)[j]
# 
#       # Find the indices in current_slice that correspond to the group_colnames
#       group_indices <- slice_colnames %in% group_colnames
# 
# 
#       # Extract the corresponding columns for the current group
#       current_group <- current_slice[, group_indices, drop = FALSE]
# 
#       # Calculate the row-wise minimum for this group
#       min_p_tensor[, j, i] <- apply(current_group, 1, min, na.rm = TRUE)
#     }
#   }
# 
#   return(min_p_tensor)
# }

```

```{r}
min_p_val <- function(p_val_tensor, gene_module_index_matrix) {
  # Get the dimensions of the input tensor
  dim_tensor <- dim(p_val_tensor)
  
  n_groups <- ncol(gene_module_index_matrix)
  
  # Calculate the number of columns per group
  cols_per_group <- dim_tensor[2] / n_groups
  
  # Initialize an empty tensor to store the results with the new dimensions
  min_p_tensor <- array(NA, dim = c(dim_tensor[1], n_groups, dim_tensor[3]),
                        dimnames = list(dimnames(p_val_tensor)[[1]], 
                                        colnames(gene_module_index_matrix), 
                                        dimnames(p_val_tensor)[[3]]))
  
  # Use mclapply to parallelize the processing over the 3rd dimension (slices)
  results <- mclapply(1:dim_tensor[3], function(i) {
    current_slice <- p_val_tensor[,,i]
    result_slice <- matrix(NA, nrow = dim_tensor[1], ncol = n_groups)
    
    # Loop over each group
    for (j in 1:n_groups) {
      # Define start and end columns for the current group
      start_col <- (j - 1) * cols_per_group + 1
      end_col <- j * cols_per_group
      
      # Ensure start_col and end_col are integers
      start_col <- as.integer(start_col)
      end_col <- as.integer(end_col)
      
      # Slice the appropriate columns for the current group
      current_group <- current_slice[, start_col:end_col, drop = FALSE]
      
      # Calculate the row-wise minimum for this group
      result_slice[, j] <- apply(current_group, 1, min, na.rm = TRUE)
    }
    
    return(result_slice)
  }, mc.cores = parallel::detectCores() - 1)  # Adjust the number of cores used
  
  # Combine the results from mclapply into the min_p_tensor
  for (i in 1:dim_tensor[3]) {
    min_p_tensor[,,i] <- results[[i]]
  }
  
  return(min_p_tensor)
}
```

```{r}
compute_final_p_val <- function(min_p_tensor, n_permute) {
  # Get the dimensions of the tensor
  dim_tensor <- dim(min_p_tensor)
  
  # Initialize an empty matrix to store the p-value counts
  final_p_vals <- matrix(0, nrow = dim_tensor[3], ncol = dim_tensor[2], 
                         dimnames = list(dimnames(min_p_tensor)[[3]], dimnames(min_p_tensor)[[2]]))
  
  # Use mclapply to parallelize over each slice of the tensor (rows)
  results <- mclapply(1:dim_tensor[3], function(i) {
    # Extract the current slice (2D matrix)
    current_slice <- min_p_tensor[,,i]
    
    # Extract the original values (the first row)
    original_values <- current_slice[1,]
    
    # Use matrix broadcasting and colSums to count the number of permutation values smaller than original
    colSums(current_slice <= matrix(original_values, nrow = nrow(current_slice), 
                                    ncol = ncol(current_slice), byrow = TRUE))
  }, mc.cores = parallel::detectCores() - 1)  # Adjust based on available cores
  
  # Assign the results to the final_p_vals matrix
  for (i in 1:dim_tensor[3]) {
    final_p_vals[i, ] <- results[[i]]
  }
  
  return(final_p_vals / (n_permute + 1))
}

compute_final_p_val_skew_normal <- function(tensor) {
  # Get the dimensions of the tensor
  dim_tensor <- dim(tensor)
  
  p_value_matrix <- matrix(NA, nrow = dim_tensor[3], ncol = dim_tensor[2])
  
  # Apply mclapply in parallel over the 3rd dimension (slices)
  result <- mclapply(1:dim_tensor[3], function(j) {
    # Apply fit_and_evaluate_skew_normal to each column vector in the 2nd dimension for the current slice
    sapply(1:dim_tensor[2], function(i) {
      fit_and_evaluate_skew_normal(tensor[,i,j], 2, T)
    })
  }, mc.cores = parallel::detectCores() - 1)  # Use all but 1 core for parallel processing
  
  # Combine results into p_value_matrix
  for (j in 1:dim_tensor[3]) {
    p_value_matrix[j,] <- result[[j]]
  }
  rownames(p_value_matrix)<-dimnames(tensor)[[3]]
  colnames(p_value_matrix)<-dimnames(tensor)[[2]]

  return(p_value_matrix)
}

fit_skew_t <- function(x,TStat1){
  x.mean = mean(x)
  x.sd = sd(x)
  x.norm = (x - x.mean)/x.sd
  mod = sgt.mle(X.f = ~x.norm,
                start = list(mu = 0, sigma = 1, lambda = 0, p = 2, q = 10))
  para = mod$estimate
  p = psgt((TStat1-x.mean)/x.sd, mu = para['mu'], sigma = para['sigma'],
           lambda = para['lambda'], p = para['p'], q = para['q']) 
  return(p)
}

compute_final_p_val_skew_normal_2 <- function(tensor) {

  dim_tensor <- dim(tensor)
  p_value_matrix <- matrix(NA, nrow = dim_tensor[3], ncol = dim_tensor[2])
  
  # Apply over the third dimension
  result <- lapply(1:dim_tensor[3], function(j) {
    sapply(1:dim_tensor[2], function(i) {
      fit_skew_t(tensor[-1, i, j],tensor[1,i,j])  # Apply function across the first dimension for each (i,j)
    })
  })

  # Combine results into p_value_matrix
  for (j in 1:dim_tensor[3]) {
    p_value_matrix[j, ] <- result[[j]]
  }

  rownames(p_value_matrix) <- dimnames(tensor)[[3]]
  colnames(p_value_matrix) <- dimnames(tensor)[[2]]

  return(p_value_matrix)
}
library(parallel)

compute_final_p_val_skew_normal_2 <- function(tensor) {

  dim_tensor <- dim(tensor)
  p_value_matrix <- matrix(NA, nrow = dim_tensor[3], ncol = dim_tensor[2])
  
  # Detect the number of cores for parallel processing
  num_cores <- detectCores() - 1  # Reserve 1 core for system processes
  
  # Parallelize the outer loop over the third dimension
  result <- mclapply(1:dim_tensor[3], function(j) {
    sapply(1:dim_tensor[2], function(i) {
      fit_skew_t(tensor[-1, i, j], tensor[1,i,j])  # Apply function across the first dimension for each (i,j)
    })
  }, mc.cores = num_cores)  # Set number of cores for parallel processing
  
  # Combine results into p_value_matrix
  for (j in 1:dim_tensor[3]) {
    p_value_matrix[j, ] <- result[[j]]
  }

  rownames(p_value_matrix) <- dimnames(tensor)[[3]]
  colnames(p_value_matrix) <- dimnames(tensor)[[2]]

  return(p_value_matrix)
}

```

```{r}
analysis_gene_module <- function (grna_assignment_matrix, response_matrix, gene_module_index_matrix, test_type, n_permute, use_resample, use_approximation) {

  func_start_time <- Sys.time()

  # permuted grna assignment matrix
  grna_assignment_matrix_combined <- permutation_or_resampling(grna_assignment_matrix, n_permute, use_resample)

  # Test statistics calculation
  step_start_time <- Sys.time()
  if (test_type == "t_test") {
    test_result <- t_test(response_matrix, grna_assignment_matrix_combined)
    step_end_time <- Sys.time()
  print(paste("Time for test statistics calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Combined statistics
  step_start_time <- Sys.time()
  combined_result <- combine_test_statistics(test_result, gene_module_index_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for combine test statistics:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Change matrix to tensor
  step_start_time <- Sys.time()
  test_tensor <- convert_matrix_to_tensor(combined_result, n_permute, grna_assignment_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for matrix to tensor conversion:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Calculate p-value
  step_start_time <- Sys.time()
  p_tensor <- compute_p_value_from_tensor(test_tensor, n_permute, 1)
    step_end_time <- Sys.time()
  print(paste("Time for p-value calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))
    

  } else if (test_type == "wilcox_test") {
    test_result <- wilcox_test(response_matrix, grna_assignment_matrix_combined)
    step_end_time <- Sys.time()
  print(paste("Time for test statistics calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Combined statistics
  step_start_time <- Sys.time()
  combined_result <- combine_test_statistics(test_result, gene_module_index_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for combine test statistics:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Change matrix to tensor
  step_start_time <- Sys.time()
  test_tensor <- convert_matrix_to_tensor(combined_result, n_permute, grna_assignment_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for matrix to tensor conversion:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Calculate p-value
  step_start_time <- Sys.time()
  p_tensor <- compute_p_value_from_tensor(test_tensor, n_permute, 2)
    step_end_time <- Sys.time()
  print(paste("Time for p-value calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))
  } else {
    stop("Unknown test type")
  }
  
  min_pval <- min_p_val(p_tensor, gene_module_index_matrix)
  
  precompute_p <- compute_final_p_val(min_pval, n_permute)
  
  if (use_approximation) {
 
    n_test <- dim(precompute_p)[1]
    threshold <- 0.1/n_test + sqrt(0.1/n_test/n_permute)
    indices <- which(precompute_p <= threshold, arr.ind = TRUE)


# Use apply to loop through indices
apply(indices, 1, function(idx) {
    i <- idx[2]
    j <- idx[1]
    
    # Perform the fit_skew_t calculation
    result <- fit_skew_t(min_pval[-1, i, j], min_pval[1, i, j])
    
    # Update final_p inside the function
    precompute_p[j, i] <<- result  
})
}
    
  

  # End timer for the entire function
func_end_time <- Sys.time()
print(paste("Total function execution time:", as.numeric(difftime(func_end_time, func_start_time, units = "secs"))))

  return(precompute_p)
}
```

## Data filtering

```{r}
setwd("/project/xuanyao/jiaming/Getting_started")
gene_names<-rownames(sceptre_object@response_matrix[[1]])

gene_modules <- readRDS("data/gene_modules_id.rds")
gene_in_modules <- vector("logical", length(gene_names))
covariate <- sceptre_object@covariate_data_frame
# Loop through each gene in gene_names
for (i in seq_along(gene_names)) {
  gene <- gene_names[i]
  
  # Check if the gene is in any module and return TRUE if it is found in any module
  gene_in_modules[i] <- any(sapply(gene_modules, function(module) gene %in% module))
}

response_matrix_raw<-sceptre_object@response_matrix[[1]][gene_in_modules,]


response_matrix_normalized<-t(t(response_matrix_raw)/covariate$response_n_umis)
response_matrix_normalized<-response_matrix_raw

row_sums <- rowSums(response_matrix_normalized!=0)

gene_to_exclude <- rownames(response_matrix_normalized)[row_sums <= 0.05*ncol(response_matrix_normalized)]

response_matrix <- response_matrix_normalized[!rownames(response_matrix_normalized) %in% gene_to_exclude,]

gene_module_index_matrix <- convert_gene_list_to_matrix(response_matrix, gene_modules)

```

```{r}
length(gene_to_exclude)
```

## Parameters

```{r}
use_resample <- T
use_approximation <- T
test_type="t_test"
n_permute <- 1000

grna_assignment_matrix <-get_grna_assignments(
  sceptre_object = sceptre_object
)
n_rows <- nrow(grna_assignment_matrix)


```

```{r}

analyze_chunk <- function(chunk) {
  analysis_gene_module(
    chunk, 
    t(rank_matrix(as.matrix(response_matrix))$X_ranked), 
    gene_module_index_matrix, 
    test_type, 
    n_permute, 
    use_resample, 
    use_approximation
  )
}

# Get the total number of rows in the grna_assignment_matrix
n_rows <- nrow(grna_assignment_matrix)

# Define the number of rows to process per iteration
chunk_size <- 30

# Split the grna_assignment_matrix into chunks of 30 rows
chunks <- split(1:n_rows, rep(1:ceiling(n_rows / chunk_size), each = chunk_size, length.out = n_rows))

# Ensure each chunk is converted back to a matrix and then apply the analysis_gene_module function
result <- do.call(rbind, lapply(chunks, function(rows) {
  chunk_matrix <- grna_assignment_matrix[rows, , drop = FALSE]  
  analyze_chunk(chunk_matrix)
}))



```

```{r}

gg_qqplot(as.vector(as.matrix(results[177:188,1:3])))
```

## Tests

```{r}
  func_start_time <- Sys.time()

  # permuted grna assignment matrix
  grna_assignment_matrix_combined <- permutation_or_resampling(grna_assignment_matrix, n_permute, use_resample)

  # Test statistics calculation
  step_start_time <- Sys.time()
  if (test_type == "t_test") {
    test_result <- t_test(response_matrix, grna_assignment_matrix_combined)
    step_end_time <- Sys.time()
  print(paste("Time for test statistics calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Combined statistics
  step_start_time <- Sys.time()
  combined_result <- combine_test_statistics(test_result, gene_module_index_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for combine test statistics:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Change matrix to tensor
  step_start_time <- Sys.time()
  test_tensor <- convert_matrix_to_tensor(combined_result, n_permute, grna_assignment_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for matrix to tensor conversion:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Calculate p-value
  step_start_time <- Sys.time()
  p_tensor <- compute_p_value_from_tensor(test_tensor, n_permute, 1)
    step_end_time <- Sys.time()
  print(paste("Time for p-value calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))
    

  } else if (test_type == "wilcox_test") {
    test_result <- wilcox_test(t(rank_matrix(as.matrix(response_matrix))$X_ranked), grna_assignment_matrix_combined)
    step_end_time <- Sys.time()
  print(paste("Time for test statistics calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Combined statistics
  step_start_time <- Sys.time()
  combined_result <- combine_test_statistics(test_result, gene_module_index_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for combine test statistics:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Change matrix to tensor
  step_start_time <- Sys.time()
  test_tensor <- convert_matrix_to_tensor(combined_result, n_permute, grna_assignment_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for matrix to tensor conversion:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Calculate p-value
  step_start_time <- Sys.time()
  p_tensor <- compute_p_value_from_tensor(test_tensor, n_permute, 2)
    step_end_time <- Sys.time()
  print(paste("Time for p-value calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))
  } else {
    stop("Unknown test type")
  }
  
  min_pval <- min_p_val(p_tensor, gene_module_index_matrix)
  
  precompute_p <- compute_final_p_val(min_pval, n_permute)
```

```{r}
wilcox_test(response_matrix, grna_assignment_matrix)[1:5,1:5]
```

```{r}
wilcox_test(t(rank_matrix(as.matrix(response_matrix))$X_ranked), grna_assignment_matrix)[1:5,1:5]
```

```{r}
wilcox.test(t(rank_matrix(as.matrix(response_matrix))$X_ranked)[2,] ~ grna_assignment_matrix[1,])
```

```{r}
class(t(rank_matrix(as.matrix(response_matrix))$X_ranked))
```

```{r}
library(ggplot2)

gg_qqplot <- function(ps, ci = 0.95) {
  n  <- length(ps)
  df <- data.frame(
    observed = -log10(sort(ps)),
    expected = -log10(ppoints(n)),
    clower   = -log10(qbeta(p = (1 - ci) / 2, shape1 = 1:n, shape2 = n:1)),
    cupper   = -log10(qbeta(p = (1 + ci) / 2, shape1 = 1:n, shape2 = n:1))
  )
  log10Pe <- expression(paste("Expected -log"[10], plain(P)))
  log10Po <- expression(paste("Observed -log"[10], plain(P)))
  ggplot(df) +
    geom_ribbon(mapping = aes(x = expected, ymin = clower, ymax = cupper),
                alpha = 0.1) +
    geom_point(aes(expected, observed), size = 1.5, color="red") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.5) +
    xlab(log10Pe) + ylab(log10Po)
}

p_new_positive <- as.vector(as.matrix(precompute_p)[1:22,])
p_new_negative <- as.vector(as.matrix(precompute_p)[23:34,])
```

```{r}
gg_qqplot(p_new_positive)
gg_qqplot(p_new_negative)

```

```{r}
  func_start_time <- Sys.time()

  # permuted grna assignment matrix
  grna_assignment_matrix_combined <- permutation_or_resampling(grna_assignment_matrix, n_permute, use_resample)

  # Test statistics calculation
  step_start_time <- Sys.time()
  if (test_type == "t_test") {
    test_result <- t_test(response_matrix, grna_assignment_matrix_combined)
    step_end_time <- Sys.time()
  print(paste("Time for test statistics calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Combined statistics
  step_start_time <- Sys.time()
  combined_result <- combine_test_statistics(test_result, gene_module_index_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for combine test statistics:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Change matrix to tensor
  step_start_time <- Sys.time()
  test_tensor <- convert_matrix_to_tensor(combined_result, n_permute, grna_assignment_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for matrix to tensor conversion:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Calculate p-value
  step_start_time <- Sys.time()
  p_tensor <- compute_p_value_from_tensor(test_tensor, n_permute, 1)
    step_end_time <- Sys.time()
  print(paste("Time for p-value calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))
    

  } else if (test_type == "wilcox_test") {
    test_result <- wilcox_test(response_matrix, grna_assignment_matrix_combined)
    step_end_time <- Sys.time()
  print(paste("Time for test statistics calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Combined statistics
  step_start_time <- Sys.time()
  combined_result <- combine_test_statistics(test_result, gene_module_index_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for combine test statistics:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Change matrix to tensor
  step_start_time <- Sys.time()
  test_tensor <- convert_matrix_to_tensor(combined_result, n_permute, grna_assignment_matrix)
  step_end_time <- Sys.time()
  print(paste("Time for matrix to tensor conversion:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))

  # Calculate p-value
  step_start_time <- Sys.time()
  p_tensor <- compute_p_value_from_tensor(test_tensor, n_permute, 2)
    step_end_time <- Sys.time()
  print(paste("Time for p-value calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))
  } else {
    stop("Unknown test type")
  }
  
  min_pval <- min_p_val(p_tensor, gene_module_index_matrix)
  
  precompute_p <- compute_final_p_val(min_pval, n_permute)
  
  if (use_approximation) {
 
    n_test <- dim(precompute_p)[1]
    threshold <- 0.1/n_test + sqrt(0.1/n_test/n_permute)
    indices <- which(precompute_p <= threshold, arr.ind = TRUE)


# Use apply to loop through indices
apply(indices, 1, function(idx) {
    i <- idx[2]
    j <- idx[1]
    
    # Perform the fit_skew_t calculation
    result <- fit_skew_t(min_pval[-1, i, j], min_pval[1, i, j])
    
    # Update final_p inside the function
    precompute_p[j, i] <<- result  
})
}

```

```{r}
sum(precompute_p<1e-3)

```

# Oct. 10

```{r}
t_test(t(as.matrix(covariate$response_n_umis)),grna_assignment_matrix)
```

```{r}
(t_test_result <- t.test(response_matrix[4,]~ grna_assignment_matrix[5,]))
```

```{r}
grna_assignment_matrix <-get_grna_assignments(
  sceptre_object = sceptre_object
)[189:200,]
response_matrix_raw<-sceptre_object@response_matrix[[1]][gene_in_modules,]


response_matrix_normalized<-response_matrix_raw


row_sums <- rowSums(response_matrix_normalized!=0)

gene_to_exclude <- rownames(response_matrix_normalized)[row_sums <= 0.05*ncol(response_matrix_normalized)]

response_matrix <- response_matrix_normalized[!rownames(response_matrix_normalized) %in% gene_to_exclude, ]
```

```{r}
use_resample <- T
use_approximation <- T
test_type="t_test"
n_permute <- 1000
grna_assignment_matrix <-get_grna_assignments(
  sceptre_object = sceptre_object
)[189:200,] # choose first 10 perturbation for demostration
n_rows <- nrow(grna_assignment_matrix)
func_start_time <- Sys.time()

  # Test statistics calculation
  step_start_time <- Sys.time()
  if (test_type == "t_test") {
    test_result <- t_test(response_matrix, grna_assignment_matrix_combined)
  } else if (test_type == "wilcox_test") {
    test_result <- wilcox_test(response_matrix, grna_assignment_matrix)
  } else {
    stop("Unknown test type")
  }
  step_end_time <- Sys.time()
  print(paste("Time for test statistics calculation:", as.numeric(difftime(step_end_time, step_start_time, units = "secs"))))
```

```{r}
t_result <- apply(t(rank_matrix(as.matrix(response_matrix))$X_ranked), 1, function(row) {
  t_test_result <- t.test(row ~ grna_assignment_matrix[1,])
  return(t_test_result$p.value)
})
```

```{r}
t_result[1:10]
```

```{r}
threshold <- 0.05/n_test + sqrt(0.05/n_test/n_permute)
threshold
```

```{r}
which(precompute_p <= threshold, arr.ind = TRUE)
```

```{r}
indices <- which(precompute_p <= threshold, arr.ind = TRUE)

# Use apply() to apply the function to each row of the indices matrix
results <- apply(indices, 1, function(idx) {
  i <- idx[2]
  j <- idx[1]
  fit_skew_t(min_pval[-1, i, j], min_pval[1, i, j])
})

# Print or process the results
results
```

```{r}
result_origin <- apply(indices, 1, function(idx) {
  i <- idx[1]
  j <- idx[2]
  precompute_p[i,j]
})
```

## Data filtering

```{r}
matrix_file_path <- "/project/xuanyao/jiaming/Getting_started/data/gene_module_author_threshold/matrix.mtx"
sparse_matrix <- readMM(matrix_file_path)
```

```{r}
library(readr)

# Read compressed files using read.table() as an alternative
barcodes <- read.table(gzfile("/project/xuanyao/jiaming/Getting_started/data/gene_module_author_threshold/barcodes.tsv.gz"), header = FALSE)


```

```{r}
features <- read.table(gzfile("/project/xuanyao/jiaming/Getting_started/data/gene_module_author_threshold/features.tsv.gz"), header = FALSE, fill = TRUE, sep = "\t")

```

```{r}
features
```

```{r}
gene_names<-features$V1
setwd("/project/xuanyao/jiaming/Getting_started")

gene_modules <- readRDS("data/gene_modules_id.rds")
gene_in_modules <- vector("logical", length(gene_names))


for (i in seq_along(gene_names)) {
  gene <- gene_names[i]
  gene_in_modules[i] <- any(sapply(gene_modules, function(module) gene %in% module))
}


```

```{r}

row_sums <- rowSums(response_matrix_normalized!=0)



features_tem <- features[gene_in_modules,]
features_reduced <- features_tem[!features_tem$V1 %in% gene_to_exclude,]
matrix_tem <- sparse_matrix[gene_in_modules,]
matrix_reduced <- matrix_tem[!features_tem$V1 %in% gene_to_exclude,]

```

```{r}
nrow(features_reduced)
nrow(matrix_reduced)
```

```{r}
nrow(features_reduced)
nrow(matrix_reduced)
```

```{r}
features_final <- rbind(features_reduced,features[36602:nrow(features),])
matrix_final <- rbind(matrix_reduced,sparse_matrix[36602:nrow(features),])
```

```{r}
write.table(features_final, file = gzfile("/project/xuanyao/jiaming/Getting_started/data/gene_module_author_threshold/features.tsv.gz"), 
            sep = "\t", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

```{r}
writeMM(matrix_final, file = "/project/xuanyao/jiaming/Getting_started/data/gene_module_author_threshold/matrix.mtx")

```

## Sceptre using min p to combine

```{r}
results_run_discovery_analysis <- readRDS("/project/xuanyao/jiaming/Getting_started/output/both_trans_author_threshold_0.05_2/results_run_discovery_analysis.rds")

```

```{r}
results_run_calibration_check <- readRDS("/project/xuanyao/jiaming/Getting_started/output/both_trans_author_threshold_0.05_2/results_run_calibration_check.rds")
```

```{r}
results_run_power_check <- readRDS("/project/xuanyao/jiaming/Getting_started/output/both_trans_author_threshold_0.05_2/results_run_power_check.rds")

```

```{r}
results_run_power_check
```

```{r}
sceptre_result <- rbind(results_run_discovery_analysis[,1:8],results_run_calibration_check[,1:8],results_run_power_check)

sceptre_result_filter <- subset(sceptre_result, pass_qc == TRUE)

```

```{r}
sceptre_result <- sceptre_result_filter[,c(1,2,3,7)]
sceptre_result
```

```{r}
sceptre_result$adjusted_p_value <- p.adjust(sceptre_result$p_value, method = "BH")

# Filter for significant results (adjusted p-value < 0.05)
significant_results_sceptre <- subset(sceptre_result, adjusted_p_value < 0.05)

# Display the significant results
significant_results_sceptre
```

```{r}

module_results <- list()

# Iterate over each response_id from results_run_discovery_analysis
for (i in seq_len(nrow(significant_results_sceptre))) {
  response_id <- significant_results_sceptre$response_id[i]
  grna_id <- significant_results_sceptre$grna_id[i]
  grna_target <- significant_results_sceptre$grna_target[i]
  pval<-significant_results_sceptre$p_value[i]
  
  # Find the modules that contain the response_id
  containing_modules <- names(gene_modules)[sapply(gene_modules, function(module) response_id %in% module)]
  
  # Store the result for this response_id and grna_id
  module_results[[i]] <- data.frame(response_id = response_id, grna_id = grna_id,grna_target=grna_target, module = containing_modules, pval=pval)
}

# Combine all results into a single data frame
module_results_df_2 <- do.call(rbind, module_results)
module_results <- module_results_df_2[,-1] 

module_results


```

```{r}
sceptre_minp <- module_results %>%
  group_by(grna_id, module) %>%
  slice_min(order_by = pval, n = 1) %>%
  ungroup()

# Display the result
sceptre_minp
```

```{r}
sum(sceptre_minp$grna_target=="non-targeting")
```

## Sceptre using ACAT to combine

```{r}
module_results <- list()

# Iterate over each response_id from results_run_discovery_analysis
for (i in seq_len(nrow(sceptre_result))) {
  response_id <- sceptre_result$response_id[i]
  grna_id <- sceptre_result$grna_id[i]
  grna_target <- sceptre_result$grna_target[i]
  pval<-sceptre_result$p_value[i]
  
  # Find the modules that contain the response_id
  containing_modules <- names(gene_modules)[sapply(gene_modules, function(module) response_id %in% module)]
  
  # Store the result for this response_id and grna_id
  module_results[[i]] <- data.frame(response_id = response_id, grna_id = grna_id,grna_target=grna_target, module = containing_modules, pval=pval)
}

# Combine all results into a single data frame
module_results_df_2 <- do.call(rbind, module_results)
module_results <- module_results_df_2[,-1] 

module_results
```

```{r}
module_results$pval[module_results$pval == 1] <- 0.999
```

```{r}
library(ACAT)
module_results_unique <- module_results %>%
  group_by(grna_id, grna_target, module) %>%
  summarize(acat_pval = ACAT(pval), .groups = "drop")

# Display the resulting dataset
module_results_unique
```

BH correction to find significant result

```{r}
module_results_unique <- module_results_unique %>%
  mutate(adjusted_pval = p.adjust(acat_pval, method = "BH"))

# Filter for significant rows (adjusted p-value < 0.05)
sceptre_ACAT <- module_results_unique %>%
  filter(adjusted_pval < 0.05)

# Display the significant results
sceptre_ACAT

```

```{r}
sum(significant_results_ACAT$grna_target=="non-targeting")
```

## Our method

```{r}
result_gene_module <- read_csv("output/t_test_on_whole_reduced_dataset_normalized_threshold.csv")
```

```{r}
result_gene_module[200,]
```

```{r}
library(ggplot2)
library(dplyr)
library(readr)
gg_qqplot <- function(ps, ci = 0.95) {
  n  <- length(ps)
  df <- data.frame(
    observed = -log10(sort(ps)),
    expected = -log10(ppoints(n)),
    clower   = -log10(qbeta(p = (1 - ci) / 2, shape1 = 1:n, shape2 = n:1)),
    cupper   = -log10(qbeta(p = (1 + ci) / 2, shape1 = 1:n, shape2 = n:1))
  )
  log10Pe <- expression(paste("Expected -log"[10], plain(P)))
  log10Po <- expression(paste("Observed -log"[10], plain(P)))
  ggplot(df) +
    geom_ribbon(mapping = aes(x = expected, ymin = clower, ymax = cupper),
                alpha = 0.1) +
    geom_point(aes(expected, observed), size = 1.5, color="red") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.5) +
    xlab(log10Pe) + ylab(log10Po)
}
gg_qqplot(as.vector(as.matrix(result_gene_module[189:198,-1])))
```

```{r}
# Step 1: Flatten the matrix of p-values (excluding the first column "X1")
p_value <- as.numeric(as.matrix(result_gene_module[,-1]))

# Step 2: Apply BH correction
p_values_adjusted <- p.adjust(p_value, method = "BH")

# Step 3: Reshape the adjusted p-values back into the original matrix shape (same as result_gene_module)
p_values_matrix <- matrix(p_values_adjusted, nrow = nrow(result_gene_module), ncol = ncol(result_gene_module) - 1)

# Step 4: Identify significant pairs (adjusted p-values < 0.05)
significant_pairs <- p_values_matrix < 0.05

# Extract the significant pairs (gene names and modules)
significant_genes <- result_gene_module$X1[row(significant_pairs)[significant_pairs]]
significant_modules <- colnames(result_gene_module)[-1][col(significant_pairs)[significant_pairs]]

# Combine significant pairs into a data frame
significant_results <- data.frame(grna = significant_genes, module = significant_modules, adjusted_p_value = p_values_matrix[significant_pairs])

# Show the significant results
significant_results

```

```{r}
significant_results_with_target <- significant_results %>%
  left_join(grna_target_data_frame, by = c("grna" = "grna_id"))

# Show the updated significant results with the grna_target column
(significant_result_ours<-significant_results_with_target[,c(1,4,2,3)])
```

```{r}
sum(significant_result_ours$grna_target=="non-targeting")	
```

```{r}
unique(significant_result_ours$grna)
```

## Use sceptre function to fit skew t

```{r}
result_gene_module <- read_csv("output/try_sceptre_skew_fit_normalized_threshold.csv")
result_gene_module[1:198,]
gg_qqplot(as.vector(as.matrix(result_gene_module[189:198,-1])))
p_value <- as.numeric(as.matrix(result_gene_module[1:198,-1]))

# Step 2: Apply BH correction
p_values_adjusted <- p.adjust(p_value, method = "BH")

# Step 3: Reshape the adjusted p-values back into the original matrix shape (same as result_gene_module)
p_values_matrix <- matrix(p_values_adjusted, nrow = 198, ncol = ncol(result_gene_module) - 1)

# Step 4: Identify significant pairs (adjusted p-values < 0.05)
significant_pairs <- p_values_matrix < 0.05

# Extract the significant pairs (gene names and modules)
significant_genes <- result_gene_module$X1[row(significant_pairs)[significant_pairs]]
significant_modules <- colnames(result_gene_module)[-1][col(significant_pairs)[significant_pairs]]

# Combine significant pairs into a data frame
significant_results <- data.frame(grna = significant_genes, module = significant_modules, adjusted_p_value = p_values_matrix[significant_pairs])

# Show the significant results
significant_results
significant_results_with_target <- significant_results %>%
  left_join(grna_target_data_frame, by = c("grna" = "grna_id"))

# Show the updated significant results with the grna_target column
(significant_result<-significant_results_with_target[,c(1,4,2,3)])
```

## Comparing 3 methods

```{r}
sceptre_minp[,c(1,3)]
sceptre_ACAT[,c(1,3)]
significant_result_ours[,c(1,3)]
```

```{r}
library(eulerr)

# Define the sets and calculate overlaps
set_minp <- unique(paste(sceptre_minp %>% pull(1), sceptre_minp %>% pull(3), sep = "_"))
set_acat <- unique(paste(sceptre_ACAT %>% pull(1), sceptre_ACAT %>% pull(3), sep = "_"))
set_ours <- unique(paste(significant_result_ours %>% pull(1), significant_result_ours %>% pull(3), sep = "_"))

# Calculate intersections manually
only_minp <- length(set_minp) - length(intersect(set_minp, set_acat)) - length(intersect(set_minp, set_ours)) + length(intersect_all)
only_acat <- length(set_acat) - length(intersect(set_acat, set_minp)) - length(intersect(set_acat, set_ours)) + length(intersect_all)
only_ours <- length(set_ours) - length(intersect(set_ours, set_minp)) - length(intersect(set_ours, set_acat)) + length(intersect_all)
minp_acat <- length(intersect_minp_acat) - length(intersect_all)
minp_ours <- length(intersect_minp_ours) - length(intersect_all)
acat_ours <- length(intersect_acat_ours) - length(intersect_all)
all_three <- length(intersect_all)

# Prepare data for eulerr
venn_data <- c(
  MinP = only_minp,
  ACAT = only_acat,
  Ours = only_ours,
  "MinP&ACAT" = minp_acat,
  "MinP&Ours" = minp_ours,
  "ACAT&Ours" = acat_ours,
  "MinP&ACAT&Ours" = all_three
)

# Plot the Venn diagram with counts
plot(euler(venn_data), 
     fills = c("red", "yellow", "green"), 
     edges = "black", 
     quantities = TRUE)  # Show quantities in the regions

```

## Only consider discovery sets (without considering control sets)

```{r}
sceptre_result <- results_run_discovery_analysis
sceptre_result_filter1 <- subset(sceptre_result, pass_qc == TRUE)
sceptre_result_filter <- subset(sceptre_result_filter1, !grepl("^CD52", grna_id))
sceptre_result <- sceptre_result_filter[,c(1,2,3,7)]
sceptre_result$adjusted_p_value <- p.adjust(sceptre_result$p_value, method = "BH")

# Filter for significant results (adjusted p-value < 0.05)
significant_results_sceptre <- subset(sceptre_result, adjusted_p_value < 0.05)

module_results <- list()

# Iterate over each response_id from results_run_discovery_analysis
for (i in seq_len(nrow(significant_results_sceptre))) {
  response_id <- significant_results_sceptre$response_id[i]
  grna_id <- significant_results_sceptre$grna_id[i]
  grna_target <- significant_results_sceptre$grna_target[i]
  pval<-significant_results_sceptre$p_value[i]
  
  # Find the modules that contain the response_id
  containing_modules <- names(gene_modules)[sapply(gene_modules, function(module) response_id %in% module)]
  
  # Store the result for this response_id and grna_id
  module_results[[i]] <- data.frame(response_id = response_id, grna_id = grna_id,grna_target=grna_target, module = containing_modules, pval=pval)
}

# Combine all results into a single data frame
module_results_df_2 <- do.call(rbind, module_results)
module_results <- module_results_df_2[,-1] 
sceptre_minp <- module_results %>%
  group_by(grna_id, module) %>%
  slice_min(order_by = pval, n = 1) %>%
  ungroup()

# Display the result
sceptre_minp

```

```{r}
module_results <- list()

# Iterate over each response_id from results_run_discovery_analysis
for (i in seq_len(nrow(sceptre_result))) {
  response_id <- sceptre_result$response_id[i]
  grna_id <- sceptre_result$grna_id[i]
  grna_target <- sceptre_result$grna_target[i]
  pval<-sceptre_result$p_value[i]
  
  # Find the modules that contain the response_id
  containing_modules <- names(gene_modules)[sapply(gene_modules, function(module) response_id %in% module)]
  
  # Store the result for this response_id and grna_id
  module_results[[i]] <- data.frame(response_id = response_id, grna_id = grna_id,grna_target=grna_target, module = containing_modules, pval=pval)
}

# Combine all results into a single data frame
module_results_df_2 <- do.call(rbind, module_results)
module_results <- module_results_df_2[,-1] 
module_results$pval[module_results$pval == 1] <- 0.999

module_results_unique <- module_results %>%
  group_by(grna_id, grna_target, module) %>%
  summarize(acat_pval = ACAT(pval), .groups = "drop")
module_results_unique <- module_results_unique %>%
  mutate(adjusted_pval = p.adjust(acat_pval, method = "BH"))

# Filter for significant rows (adjusted p-value < 0.05)
sceptre_ACAT <- module_results_unique %>%
  filter(adjusted_pval < 0.05)

# Display the significant results
sceptre_ACAT
```

```{r}
n_analyze =178
p_value <- as.numeric(as.matrix(result_gene_module[1:n_analyze,-1]))

# Step 2: Apply BH correction
p_values_adjusted <- p.adjust(p_value, method = "BH")

# Step 3: Reshape the adjusted p-values back into the original matrix shape (same as result_gene_module)
p_values_matrix <- matrix(p_values_adjusted, nrow = n_analyze, ncol = ncol(result_gene_module) - 1)

# Step 4: Identify significant pairs (adjusted p-values < 0.05)
significant_pairs <- p_values_matrix < 0.05

# Extract the significant pairs (gene names and modules)
significant_genes <- result_gene_module$X1[row(significant_pairs)[significant_pairs]]
significant_modules <- colnames(result_gene_module)[-1][col(significant_pairs)[significant_pairs]]

# Combine significant pairs into a data frame
significant_results <- data.frame(grna = significant_genes, module = significant_modules, adjusted_p_value = p_values_matrix[significant_pairs])

# Show the significant results
significant_results_with_target <- significant_results %>%
  left_join(grna_target_data_frame, by = c("grna" = "grna_id"))

# Show the updated significant results with the grna_target column
(significant_result_ours<-significant_results_with_target[,c(1,4,2,3)])
```

```{r}
set_minp <- unique(paste(sceptre_minp %>% pull(1), sceptre_minp %>% pull(3), sep = "_"))
set_acat <- unique(paste(sceptre_ACAT %>% pull(1), sceptre_ACAT %>% pull(3), sep = "_"))
set_ours <- unique(paste(significant_result_ours %>% pull(1), significant_result_ours %>% pull(3), sep = "_"))

# Calculate intersections manually
only_minp <- length(set_minp) - length(intersect(set_minp, set_acat)) - length(intersect(set_minp, set_ours)) + length(intersect_all)
only_acat <- length(set_acat) - length(intersect(set_acat, set_minp)) - length(intersect(set_acat, set_ours)) + length(intersect_all)
only_ours <- length(set_ours) - length(intersect(set_ours, set_minp)) - length(intersect(set_ours, set_acat)) + length(intersect_all)
minp_acat <- length(intersect_minp_acat) - length(intersect_all)
minp_ours <- length(intersect_minp_ours) - length(intersect_all)
acat_ours <- length(intersect_acat_ours) - length(intersect_all)
all_three <- length(intersect_all)

# Prepare data for eulerr
venn_data <- c(
  MinP = only_minp,
  ACAT = only_acat,
  Ours = only_ours,
  "MinP&ACAT" = minp_acat,
  "MinP&Ours" = minp_ours,
  "ACAT&Ours" = acat_ours,
  "MinP&ACAT&Ours" = all_three
)

# Plot the Venn diagram with counts
plot(euler(venn_data), 
     fills = c("red", "yellow", "green"), 
     edges = "black", 
     quantities = TRUE)  # Show quantities in the regions
```

## Test on gene p value

```{r}
t_test_on_whole_dataset_gene_wise <- read_csv("output/t_test_on_whole_dataset_gene_wise.csv")
```

```{r}
ncol(t_test_on_whole_dataset_gene_wise)
```

```{r}
results_run_discovery_analysis[,c(1,2,7)]
```

```{r}
t_test_on_whole_dataset_gene_wise[1:5,1:5]
```

```{r}
length(unique(results_run_discovery_analysis$response_id))
```

```{r}
colnames(t_test_on_whole_dataset_gene_wise) <- c("grna",rownames(response_matrix))
```

```{r}
t_test_on_whole_dataset_gene_wise[,1:5]
```

```{r}
results_run_discovery_analysis[,c(1,2,7)]
```

```{r}
selected_rows <- results_run_discovery_analysis[results_run_discovery_analysis$response_id %in% rownames(response_matrix), c(1,2,7)]
```

```{r}
selected_rows
```

```{r}
matrix_data <- dcast(selected_rows, grna_id ~ response_id, value.var = "p_value")
```

```{r}
matrix_data_2<-matrix_data[,-1]
rownames(matrix_data_2)<-matrix_data[,1]
```

```{r}
matrix_data_2[1:5,1:5]
```

```{r}
our_test<-as.matrix(t_test_on_whole_dataset_gene_wise[1:178,-1])
rownames(our_test) <- rownames(grna_assignment_matrix)[1:178]
```

```{r}
dim(matrix_data_2)
```

```{r}
dim(our_test)
```

```{r}
our_test[1:5,1:3]
```

```{r}
sum(is.na(matrix_data_2))
```

```{r}
matrix_data_ordered <- matrix_data_2[match(rownames(our_test), rownames(matrix_data_2)), ]

# Reorder the columns based on t_test_on_whole_dataset_gene_wise column order
matrix_data_ordered <- matrix_data_ordered[, match(colnames(our_test), colnames(matrix_data_ordered))]
```

```{r}
sceptre_result <- matrix_data_ordered
```

```{r}
sceptre_result[1:5,1:3]
```

```{r}
sum(our_test<1e-30)
```

```{r}
vect1 <- as.numeric(as.matrix(sceptre_result))
vect2 <- as.numeric(as.matrix(our_test))

# Filter out NA values
valid_indices <- which(!is.na(vect1) & !is.na(vect2) )

# Create filtered vectors without NAs
filtered_vect1 <- vect1[valid_indices]
filtered_vect2 <- vect2[valid_indices]
```

```{r}
plot(filtered_vect1,filtered_vect2)
```

```{r}
t_test_on_whole_dataset_gene_wise[179:188,1:3]
```

```{r}
t_test_on_whole_dataset_gene_wise[which(t_test_on_whole_dataset_gene_wise$grna=="SNP-10_1"),"ENSG00000001167"]
```

```{r}
covariate
```

```{r}
umis_quantiles <- quantile(covariate$response_n_umis, probs = c(0.01, 0.99))
nonzero_quantiles <- quantile(covariate$response_n_nonzero, probs = c(0.01, 0.99))

# Identify rows that are outside the quantile ranges
out_of_range_indices <- which(
  covariate$response_n_umis < umis_quantiles[1] | 
  covariate$response_n_umis > umis_quantiles[2] | 
  covariate$response_n_nonzero < nonzero_quantiles[1] | 
  covariate$response_n_nonzero > nonzero_quantiles[2]
)

# Display the indices
out_of_range_indices
```
