---
title: "newest results"
author: "jliucx"
date: "2025-6-1"
output: 
  workflowr::wflow_html:
    number_sections: true
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72

---
```{r,include=F}

setwd("/project/xuanyao/jiaming/Getting_started")
Sys.setenv(TMPDIR = "tmp")
library(ggplot2)
library(dplyr)
library(readr)
library(tibble)
library(tidyr)
library(presto)
library(sceptre)
library(sceptredata)
library(readr)
library(dplyr)
library(Matrix)
library(parallel)
library(BH)
library(Rcpp)
library(peakRAM)
library(sgt)
library(heavytailcombtest)
library(gridExtra)
library(eulerr)
library(ggplot2)
library(dplyr)
library(readr)
library(tibble)
library(ACAT)
library(MASS)
source("code/functions_optimized.R")

library(tidyverse)

gg_qqplot_list <- function(pval_list, ci = 0.95, cap = 6, title = "QQ Plot") {
  # pval_list: named list of numeric p-value vectors

  # 1. 构造数据框：method, p
  pts_df <- map_dfr(names(pval_list), function(name) {
    tibble(
      Method = name,
      p = pval_list[[name]]
    )
  })

  # 2. 清洗和排序
  pts_df <- pts_df %>%
    filter(!is.na(p), p > 0) %>%
    mutate(
      observed = pmin(-log10(p), cap)
    ) %>%
    group_by(Method) %>%
    arrange(p, .by_group = TRUE) %>%
    mutate(
      rank_idx = row_number(),
      n = n(),
      expected = -log10(ppoints(n))[rank_idx]
    ) %>%
    ungroup()

  # 3. 构造 envelope（以最大 n 为准）
  max_n <- max(pts_df$n)
  exp_env <- -log10(ppoints(max_n))
  lower_env <- pmin(-log10(qbeta((1 - ci)/2, seq_len(max_n), rev(seq_len(max_n)))), cap)
  upper_env <- pmin(-log10(qbeta((1 + ci)/2, seq_len(max_n), rev(seq_len(max_n)))), cap)

  env_df <- tibble(
    expected = c(exp_env, rev(exp_env)),
    ymin = c(lower_env, rev(upper_env)),
    ymax = c(upper_env, rev(lower_env))
  )

  # 4. 作图
  ggplot() +
    geom_ribbon(data = env_df, aes(x = expected, ymin = ymin, ymax = ymax),
                fill = "grey80", alpha = 0.3) +
    geom_line(data = pts_df,
              aes(x = expected, y = observed, color = Method, group = Method),
              size = 1) +
    geom_point(data = pts_df,
               aes(x = expected, y = observed, color = Method),
               size = 1.5, alpha = 0.6) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", alpha = 0.5) +
    labs(
      title = title,
      x = expression(Expected~-log[10](italic(P))),
      y = expression(Observed~-log[10](italic(P))),
      color = "Method"
    ) +
    ylim(0, cap) +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}


gg_qqplot <- function(ps, ci = 0.95, title = "QQ Plot") {
  n <- length(ps)
  df <- data.frame(
    observed = pmin(-log10(sort(ps)), 6), # Cap observed values at 4
    expected = -log10(ppoints(n)),
    clower = pmin(-log10(qbeta(p = (1 - ci) / 2, shape1 = 1:n, shape2 = n:1)), 6), # Cap clower values at 4
    cupper = pmin(-log10(qbeta(p = (1 + ci) / 2, shape1 = 1:n, shape2 = n:1)), 6)  # Cap cupper values at 4
  )

  log10Pe <- expression(paste("Expected -log"[10], plain(P)))
  log10Po <- expression(paste("Observed -log"[10], plain(P)))

  ggplot(df) +
    geom_ribbon(mapping = aes(x = expected, ymin = clower, ymax = cupper), alpha = 0.1) +
    geom_point(aes(expected, observed), size = 1.5, color = "red") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.5) +
    xlab(log10Pe) + 
    ylab(log10Po) +
    ylim(0, 6) + # Limit y-axis to 0-5
    ggtitle(title) # Add title
}



library(ggplot2)
library(dplyr)
library(tidyr)

gg_qqplot_df <- function(pmat,
                         ci    = 0.95,
                         title = "QQ Plot",
                         cap   = 6) {
  # pmat: numeric data.frame or matrix; rows = methods, cols = p-values
  # ci:   confidence level for the envelope
  # title: plot title
  # cap:  maximum -log10(p) to display
  
  # 1. reshape & filter out NAs, compute expected & observed per method
  pmat[pmat<=0]<-1e-250
    pmat[pmat>=1]<-1-1e-250
  pts_df <- pmat %>% 
    as.data.frame() %>% 
    rownames_to_column("Method") %>% 
    pivot_longer(-Method, names_to="dummy", values_to="p") %>% 
    filter(!is.na(p)) %>%          # drop NAs
    group_by(Method) %>% 
    arrange(p) %>% 
    mutate(
      rank_idx = row_number(),
      n_i      = n(),                                 # how many p’s
      expected = -log10(ppoints(n_i))[rank_idx],      # expected quantile
      observed = pmin(-log10(p), cap)                 # cap at ‘cap’
    ) %>% 
    ungroup()
  
  # 2. build a single envelope for the *largest* group size
  max_n <- pts_df %>% pull(n_i) %>% max()
  exp_env <- -log10(ppoints(max_n))
  lower_env <- pmin(-log10(qbeta((1-ci)/2,  seq_len(max_n), rev(seq_len(max_n)))), cap)
  upper_env <- pmin(-log10(qbeta((1+ci)/2,  seq_len(max_n), rev(seq_len(max_n)))), cap)
  env_df <- tibble(expected = exp_env, ymin = lower_env, ymax = upper_env)
  
  # 3. plot
  ggplot() +
    geom_ribbon(data = env_df,
                aes(x = expected, ymin = ymin, ymax = ymax),
                fill  = "grey80", alpha = 0.3) +
    geom_line(data = pts_df,
              aes(x = expected, y = observed, color = Method, group = Method),
              size = 1) +
    geom_point(data = pts_df,
               aes(x = expected, y = observed, color = Method),
               size = 1.5) +
    geom_abline(intercept = 0, slope = 1,
                linetype = "dashed", alpha = 0.5) +
    labs(title = title,
         x     = expression(Expected~ -log[10](P)),
         y     = expression(Observed~ -log[10](P)),
         color = "Method") +
    ylim(0, cap) +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}



```


# Morris - HALLMARK

<span style="color: red;">I realized a bug in previous code: I didn't correctly normalize the expression matrix. Therefore, I rerun all the dataset.</span>


## Strong perturbation - gene qc

```{r,echo=F,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_ACAT_Morris_Hallmark_strong_gene_qc/p_values/geneset_p_value.rds"))


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc_complete_gene.rds")
 


sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  
  print(length(matched_targets))
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_normal_morris_HALLMARK_strong_gene_qc/p_values/geneset_p_value.rds"))

max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")


max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_morris_HALLMARK_strong_gene_qc/p_values/geneset_p_value.rds"))

max_stat[max_stat==0]<- 1e-4
max_stat_long_3 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3 <- max_stat_long_3 %>% filter(!is.na(p_value))


max_stat_long_3 <- max_stat_long_3 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_3$bh_pval <-  p.adjust(max_stat_long_3$p_value, method = "BH")
max_stat_long_3$identifier <- paste(max_stat_long_3$grna_id, max_stat_long_3$module, sep = "_")


manually_created_gene_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/manually_created_gene_index_matrix.rds")
```


```{r,echo=F,include=F}
sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated_full_NTC/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]


sceptre_wide <- sceptre_wide[, rownames(manually_created_gene_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(manually_created_gene_index_matrix)) {
    
    
    module_name <- colnames(manually_created_gene_index_matrix)[j]  # Get the module name

    module_genes <- rownames(manually_created_gene_index_matrix)[manually_created_gene_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]


```

```{r,echo=F}
 genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_morris_HALLMARK_strong_gene_qc/p_values/genewise_p_value.rds"))

genewise_p_combined_all[genewise_p_combined_all>=1]<- 1-1e-5
genewise_p_combined_all[genewise_p_combined_all==0]<- 1e-30
 
module_results <- list()

for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t_3000<- do.call(rbind, module_results)

module_results_t_3000 <- module_results_t_3000[!is.na(module_results_t_3000$acat_pval), ]

module_results_t_NTC_3000 <- module_results_t_3000 %>%
  filter(grepl("^NTC", grna_id))


module_results_t_3000 <- module_results_t_3000 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_t_3000$bh_pval <- p.adjust(module_results_t_3000$acat_pval, method = "BH")
module_results_t_3000$identifier <- paste(module_results_t_3000$grna_id, module_results_t_3000$module, sep = "_")
```



```{r,cache=F,fig.width=16,fig.height=8,echo=F}
library(eulerr)

significant_levels <- c(0.05,0.1)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_gene_wise_t_ACAT <- na.omit(
  module_results_sceptre$identifier[module_results_t_3000$bh_pval < significant_level]
)

significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)

significant_ids_max_stat_3 <- na.omit(
  max_stat_long_3$identifier[max_stat_long_3$bh_pval < significant_level]
)
  
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT` = significant_ids_gene_wise_p,
    `Power + ACAT`=significant_ids_max_stat,
    `Our t`=significant_ids_max_stat_3,
    `wilcoxon 3000 (t)+ACAT`=significant_ids_gene_wise_t_ACAT
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 2)

par(mfrow = c(1, 1))
```


## NTC

```{r,fig.width=15,fig.height=12,cache=F}

p6<-gg_qqplot(as.vector(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/10000_no_morris_HALLMARK_strong_gene_qc_NTC/p_values/geneset_p_value.rds")[c(1:9,12),]),title="10000 no approximation")
p7<-gg_qqplot(as.vector(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_morris_HALLMARK_strong_gene_qc_NTC/p_values/geneset_p_value.rds")[1:9,]), title="3000 t approximation (remove NTC-3, NTC-11)")

p8<-gg_qqplot(module_results_sceptre_NTC[module_results_sceptre_NTC$grna_id!="NTC-3"&module_results_sceptre_NTC$grna_id!="NTC-11",]$acat_pval, title="Sceptre+ACAT")

genewise_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_morris_HALLMARK_strong_gene_qc_NTC_p2/p_values/genewise_p_value.rds")
p9<-gg_qqplot_df((genewise_p_value), title="wilcoxon NTC genewise p")

p1<-gg_qqplot_df((readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_morris_HALLMARK_strong_gene_qc_NTC_p2/p_values/geneset_p_value.rds")),title="Negative control (correct normalization)")
p2<-gg_qqplot_df((readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_morris_HALLMARK_strong_gene_qc_NTC_previous/p_values/geneset_p_value.rds")),title="Negative control (wrong normalization)")
grid.arrange(p6,p7,p8,p9,p1,p2,ncol=2)
```

```{r,fig.width=15,fig.height=12,cache=F}

p6<-gg_qqplot(as.vector(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_ACAT_Morris_Hallmark_strong_gene_qc/p_values/geneset_p_value.rds")[c(189:191,192,193:197,208),]),title="power combination + ACAT (Remove NTC 3, 11)")
p7<-gg_qqplot(as.vector(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_ACAT_Morris_Hallmark_strong_gene_qc/p_values/geneset_p_value.rds")[c(189:191,193:197,208),]),title="power combination + ACAT (Remove NTC 3, 5, 11)")
p8 <- gg_qqplot_df(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Morris_Hallmark/3000_t_ACAT_Morris_Hallmark_strong_gene_qc/p_values/geneset_p_value.rds")[c(189:198,207,208),],title = "power combination + ACAT (All NTC)")
grid.arrange(p6,p7,p8,ncol=2)
```




# Xie - HALLMARK



## Strong perturbation - gene qc

```{r,echo=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/program_data/gene_module_index_matrix.rds")


grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/program_data/grna_target_data_frame.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/program_data/discovery_pairs_cis.rds")


max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Xie_Hallmark/3000_t_Xie_HALLMARK_strong_gene_qc/p_values/geneset_p_value.rds"))


sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Xie/sceptre/adj_crt_union_batch_threshold_4/results_run_discovery_analysis.rds")[,c(1,2,6)]
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/Xie_perturbation_gene_qc.rds")
 

gene_names <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/program_data/gene_to_analyze_filtered.rds")

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_target
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, gene_names, drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_id
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}

  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- gene_names[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

```


```{r,echo=F}
max_stat_2 <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Xie_Hallmark/10000_no_Xie_HALLMARK_strong_gene_qc/p_values/geneset_p_value.rds"))

max_stat_long_2 <- max_stat_2%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))






max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))




max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")


max_stat_3 <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Xie_Hallmark/3000_t_ACAT_Xie_Hallmark_strong_gene_qc/p_values/geneset_p_value.rds"))

max_stat_long_3 <- max_stat_3%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3 <- max_stat_long_3 %>% filter(!is.na(p_value))






max_stat_long_3 <- max_stat_long_3 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))




max_stat_long_3$bh_pval <-  p.adjust(max_stat_long_3$p_value, method = "BH")
max_stat_long_3$identifier <- paste(max_stat_long_3$grna_id, max_stat_long_3$module, sep = "_")
```






```{r,cache=F,fig.width=16,fig.height=8,echo=F}
library(eulerr)

significant_levels <- c(0.05,0.1)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)



significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)

significant_ids_max_stat_3 <- na.omit(
  max_stat_long_3$identifier[max_stat_long_3$bh_pval < significant_level]
)
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT` = significant_ids_gene_wise_p,
    `Our t`=significant_ids_max_stat,
    `Power comb + ACAT`=significant_ids_max_stat_3
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 2)

par(mfrow = c(1, 1))
```


## NTC

```{r,cache=F,fig.width=15,fig.height=10}
p1<-gg_qqplot(as.vector(as.matrix(max_stat[519,])),title="Gene module p value NTC-1 (our method t)")
p2<-gg_qqplot(as.vector(as.matrix(max_stat[521,])),title="Gene module p value NTC-3 (our method t)")


p3<-gg_qqplot(as.vector(as.matrix(max_stat_3[519,])),title="Gene module p value NTC-1 (power comb + ACAT)")
p4<-gg_qqplot(as.vector(as.matrix(max_stat_3[521,])),title="Gene module p value NTC-3 (power comb + ACAT)")

grid.arrange(p1,p2,p3,p4,ncol=2)
```

# Xie - cNMF set

```{r,echo=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/program_data_cNMF/gene_module_index_matrix.rds")


grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/program_data_cNMF/grna_target_data_frame.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/program_data_cNMF/discovery_pairs_cis.rds")


max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Xie_cNMF/3000_t_Xie_cNMF_strong_gene_qc/p_values/geneset_p_value.rds"))


sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Xie/sceptre/cNMF_threshold4/results_run_discovery_analysis.rds")[,c(1,2,6)]
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Xie/Xie_perturbation_gene_qc.rds")
 

gene_names <- rownames(gene_module_index_matrix)

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_target
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, gene_names, drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_id
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}

  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- gene_names[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

```

```{r}
max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/Xie_cNMF/3000_t_ACAT_Xie_cNMF_strong_gene_qc/p_values/geneset_p_value.rds"))

max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")

```



```{r,cache=F,fig.width=16,fig.height=8,echo=F}
library(eulerr)

significant_levels <- c(0.05,0.1)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)



significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)
  venn_data <- list(
    `Sceptre+ACAT` = significant_ids_gene_wise_p,
    `Our t`=significant_ids_max_stat,
    `Power combine + ACAT` = significant_ids_max_stat_2
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 2)

par(mfrow = c(1, 1))
```

# Repologle

## Hallmark set

### maxstat

```{r,fig.width=60,fig.height=25}
geneset_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/repologle/NTC_maxstat_10000_no/p_values/geneset_p_value.rds")
combined_mat<-geneset_p_value 
ind <- which(matrixStats::rowMins(as.matrix(combined_mat)) < 0.0007)
rownames(combined_mat)<-paste0(1:nrow(combined_mat))
gg_qqplot_df(combined_mat,title="Power + Maxstat (global permutation)")
```
```{r,fig.width=60,fig.height=25}
geneset_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/repologle/NTC_within_batch_permute_maxstat_10000_no/p_values/geneset_p_value.rds")
combined_mat<-geneset_p_value 
rownames(combined_mat)<-paste0(1:nrow(combined_mat))
gg_qqplot_df(combined_mat)
```




### powersum + acat

```{r,fig.width=60,fig.height=25}
geneset_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/repologle/NTC_power_ACAT_10000_no/p_values/geneset_p_value.rds")
combined_mat<-geneset_p_value 
ind <- which(matrixStats::rowMins(as.matrix(combined_mat)) < 0.0007)
rownames(combined_mat)<-paste0(1:nrow(combined_mat))
gg_qqplot_df(combined_mat)
```


### sceptre genewise




```{r,echo=F}
results_run_calibration_check <- readRDS("/project/xuanyao/jiaming/Getting_started/output/repologle/permutation_complete_NTC/results_run_calibration_check.rds")
grna_groups <- split(results_run_calibration_check$p_value, results_run_calibration_check$grna_id)
```


```{r,fig.width=60,fig.height=25}
names(grna_groups)<-1:length(grna_groups)
gg_qqplot_list(grna_groups,title="Sceptre: genewise NTC")
```


# Papalexi

## permutations

<span style="color: red;"> For some details of obtaining those results:
Both sceptre and I used permutation, instead of conditional resampling, as it is a low moi dataset.
And for this dataset, the cell covariate contains many categorical variables, i.e. phases, bio_rep, etc. So I wonder whether glm is better than wilcoxon in terms of handling confounding in this case, when both methods use permutation instead of conditional resampling to obtain p values. </span>

```{r}
cell_covariate <- readRDS("/project/xuanyao/jiaming/Getting_started/data/papalexi/cell_covariate.rds")
head(cell_covariate)
```


```{r,echo=F,include=F}
gene_module_index_matrix <-  readRDS("/project/xuanyao/jiaming/Getting_started/data/papalexi/gene_module_index_matrix.rds")

grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/papalexi/grna_target_dataframe.rds")

discovery_pairs_cis <-readRDS("/project/xuanyao/jiaming/Getting_started/data/papalexi/discovery_pairs_cis.rds")


max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/10000_no_ACAT/p_values/geneset_p_value.rds"))


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/papalexi/hallmark_set_genes/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- character(0) 
 


sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, colnames(sceptre_wide)%in%rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]
```



```{r,echo=F,include=F}
library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/10000_no_maxstat/p_values/geneset_p_value.rds"))

max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")
```




```{r,cache=F,fig.width=16,fig.height=8,echo=F}
library(eulerr)

significant_levels <- c(0.1)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots

significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_levels]
)


significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_levels]
)
significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_levels]
)
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT` = significant_ids_gene_wise_p,
    `Power + ACAT`=significant_ids_max_stat,
    `Power + Maxstat`=significant_ids_max_stat_2
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot


```


```{r,include=F}
geneset_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/10000_no_ACAT/p_values/geneset_p_value.rds")

 gene_within_hallmark_geneset <- readRDS("/project/xuanyao/jiaming/Getting_started/data/papalexi/gene_within_hallmark_geneset.rds")
p1<-gg_qqplot_df(geneset_p_value[which(grepl("NTg",rownames(geneset_p_value))),],title="wilcoxon geneset NTC")
 results_run_calibration_check <- readRDS("/project/xuanyao/jiaming/Getting_started/output/papalexi/hallmark_set_genes/results_run_calibration_check.rds")
results_run_calibration_check <- results_run_calibration_check[results_run_calibration_check$response_id%in%gene_within_hallmark_geneset ,]

# 假设数据名为 results_run_calibration_check
grna_pval_list <- split(results_run_calibration_check$p_value, results_run_calibration_check$grna_id)


p2<-gg_qqplot_list(grna_pval_list,title="sceptre genewise NTC")

genewise_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/10000_no_maxstat/p_values/genewise_p_value.rds")
p3<-gg_qqplot_df(genewise_p_value[which(grepl("NTg",rownames(genewise_p_value))),],title="wilcoxon genewise NTC")
```



```r

df1 <- as.data.frame(genewise_p_value)  # 106 x 3725
df2 <- sceptre_wide                    # 97 x 3334

common_rows <- intersect(rownames(df1), rownames(df2))
common_cols <- intersect(colnames(df1), colnames(df2))


df1_common <- df1[common_rows, common_cols]
df2_common <- df2[common_rows, common_cols]

df1_long <- as.vector(as.matrix(df1_common))
df2_long <- as.vector(as.matrix(df2_common))

scatter_df <- data.frame(df1 = df1_long, df2 = df2_long)
scatter_df <- na.omit(scatter_df)
library(ggplot2)

ggplot(scatter_df, aes(x = df1, y = df2)) +
  geom_point(alpha = 0.5) +
  theme_minimal() +
  labs(x = "genewise_p_value", y = "sceptre_wide", title = "Scatter Plot of Common Entries")


```





```{r,fig.height=8,fig.width=12}
grid.arrange(euler_plot,p1,p3,p2,ncol=2)
```

## resample

<span style="color: red;"> I tried conditional resampling instead of permutations, and the inflation problem is solved.
 </span>

```{r,echo=F,include=F}
gene_module_index_matrix <-  readRDS("/project/xuanyao/jiaming/Getting_started/data/papalexi/gene_module_index_matrix.rds")

grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/papalexi/grna_target_dataframe.rds")

discovery_pairs_cis <-readRDS("/project/xuanyao/jiaming/Getting_started/data/papalexi/discovery_pairs_cis.rds")


max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/3000_t_ACAT_resample/p_values/geneset_p_value.rds"))


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/papalexi/hallmark_set_genes/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- character(0) 
 


sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, colnames(sceptre_wide)%in%rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]
```



```{r,echo=F,include=F}
library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/3000_t_maxstat_resample/p_values/geneset_p_value.rds"))

max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")
```

```{r,include=F}
 genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/3000_t_ACAT_resample/p_values/genewise_p_value.rds"))

genewise_p_combined_all[genewise_p_combined_all>=1]<- 1-1e-5
genewise_p_combined_all[genewise_p_combined_all==0]<- 1e-30
 
module_results <- list()

for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t_3000<- do.call(rbind, module_results)

module_results_t_3000 <- module_results_t_3000[!is.na(module_results_t_3000$acat_pval), ]

module_results_t_NTC_3000 <- module_results_t_3000 %>%
  filter(grepl("^NTC", grna_id))


module_results_t_3000 <- module_results_t_3000 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_t_3000$bh_pval <- p.adjust(module_results_t_3000$acat_pval, method = "BH")
module_results_t_3000$identifier <- paste(module_results_t_3000$grna_id, module_results_t_3000$module, sep = "_")
```



```{r,cache=F,fig.width=16,fig.height=8,echo=F}
library(eulerr)

significant_levels <- c(0.1)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots

significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_levels]
)


significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_levels]
)
significant_ids_max_stat_2 <- na.omit(
  module_results_t_3000$identifier[module_results_t_3000$bh_pval < significant_levels]
)
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT` = significant_ids_gene_wise_p,
    `Power + ACAT`=significant_ids_max_stat,
    `Wilcoxon + ACAT`=significant_ids_max_stat_2
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_levels)
)
plot_list[[length(plot_list) + 1]] <- euler_plot


```


```{r,include=F}
geneset_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/3000_t_ACAT_resample/p_values/geneset_p_value.rds")
p1<-gg_qqplot_df(geneset_p_value[which(grepl("NTg",rownames(geneset_p_value))),],title="wilcoxon geneset NTC")

results_run_calibration_check <- results_run_calibration_check[results_run_calibration_check$response_id%in%gene_within_hallmark_geneset ,]

# 假设数据名为 results_run_calibration_check
grna_pval_list <- split(results_run_calibration_check$p_value, results_run_calibration_check$grna_id)


p2<-gg_qqplot_list(grna_pval_list,title="sceptre genewise NTC")

genewise_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/papalexi/3000_t_ACAT_resample/p_values/genewise_p_value.rds")
p3<-gg_qqplot_df(genewise_p_value[which(grepl("NTg",rownames(genewise_p_value))),],title="wilcoxon genewise NTC")
```







```{r,fig.height=8,fig.width=12}
grid.arrange(euler_plot,p1,p3,p2,ncol=2)
```

# Frangieh - IFN gamma


<span style="color: red;"> It is worth noticing that different from other datasets, this dataset contains cell-cycle related information: s_score, g2m_score, and phase (according to the author, phase is calculated using s_score and g2m_score). In sceptre paper, the author take those additional covariate in their glm model, and use permutation to get p value.</span>

<span style="color: red;"> To imitate this, I also take this three additional covariate, and (use conditional resampling to obtain p values.) </span>

```{r}
cell_covariate <- readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/ifn_gamma/cell_covariate.rds")

head(cell_covariate )
```

```{r,echo=F,include=F}
gene_module_index_matrix <-  readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/ifn_gamma/hallmark_set/gene_module_index_matrix.rds")

grna_target_data_frame_omit <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/ifn_gamma/grna_target_dataframe.rds"))

discovery_pairs_cis <-readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/ifn_gamma/discovery_pairs_cis.rds")


max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh/resample_ACAT/p_values/geneset_p_value.rds"))


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/frangieh/hallmark_set_genes/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/ifn_gamma/hallmark_set/perturbation_gene_qc.rds")
 


sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, colnames(sceptre_wide)%in%rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]
```





```{r,echo=F,include=F}
library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id == grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
  print(length(remove_genes))
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
  print("no genes removed")
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh/resample_maxstat/p_values/geneset_p_value.rds"))

max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")


max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh/resample_ACAT_50_qc/p_values/geneset_p_value.rds"))

max_stat[max_stat==0]<- 1e-4
max_stat_long_3 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3 <- max_stat_long_3 %>% filter(!is.na(p_value))


max_stat_long_3 <- max_stat_long_3 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_3$bh_pval <-  p.adjust(max_stat_long_3$p_value, method = "BH")
max_stat_long_3$identifier <- paste(max_stat_long_3$grna_id, max_stat_long_3$module, sep = "_")
```









```{r,include=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots

significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_levels]
)


significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_levels]
)
significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_levels]
)

significant_ids_max_stat_3 <- na.omit(
  max_stat_long_3$identifier[max_stat_long_3$bh_pval < significant_levels]
)
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT (strong qc)` = significant_ids_gene_wise_p,
    `Power + ACAT (default qc)`=significant_ids_max_stat,
 `Power + ACAT (strong qc)`=significant_ids_max_stat_3
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_levels)
)
p1 <- euler_plot


```




```{r,include=F}
sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/frangieh/hallmark_set_genes/results_run_calibration_check.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/ifn_gamma/hallmark_set/perturbation_gene_qc.rds")
 


sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, colnames(sceptre_wide)%in%rownames(gene_module_index_matrix), drop = FALSE]
sceptre_wide<-sceptre_wide  [grepl("NO_SITE",rownames(sceptre_wide )),]

module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id == grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]
```


```{r,include=F}
sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/frangieh/hallmark_set_genes/results_run_calibration_check.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- character(0)
 


sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, colnames(sceptre_wide)%in%rownames(gene_module_index_matrix), drop = FALSE]
sceptre_wide<-sceptre_wide  [grepl("NO_SITE",rownames(sceptre_wide )),]

module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id == grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC_default <- do.call(rbind, module_results)
module_results_sceptre_NTC_default <- module_results_sceptre_NTC_default[!is.na(module_results_sceptre_NTC_default$acat_pval), ]
```



```{r,include=F}
genewise_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh/resample_ACAT/p_values/genewise_p_value.rds")

build_overlap_lists <- function(A, B) {
  # 行、列只取公共部分
  common_rows <- intersect(rownames(A), rownames(B))
  common_cols <- intersect(colnames(A), colnames(B))
  A <- A[common_rows, common_cols, drop = FALSE]
  B <- B[common_rows, common_cols, drop = FALSE]

  gene_list   <- vector("list", length(common_rows))
  sceptre_list<- vector("list", length(common_rows))
  names(gene_list) <- names(sceptre_list) <- common_rows

  for (r in common_rows) {
    ok <- !is.na(A[r, ]) & !is.na(B[r, ])      # 两边都非 NA
    if (any(ok)) {
      gene_list[[r]]    <- as.numeric(A[r, ok])
      sceptre_list[[r]] <- as.numeric(B[r, ok])
    } else {
      gene_list[[r]]    <- numeric(0)
      sceptre_list[[r]] <- numeric(0)
    }
  }
  list(genewise = gene_list, sceptre = sceptre_list)
}
```


```{r,include=F}
p2 <- gg_qqplot_list(build_overlap_lists(genewise_p_value,sceptre_wide)[["genewise"]]
,title="GENEWISE : Wilcoxon")
p3 <- gg_qqplot_list(build_overlap_lists(genewise_p_value,sceptre_wide)[["sceptre"]]
,title = "GENEWISE : Sceptre")
```





```{r,include=F}
geneset_p_value <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh/resample_ACAT/p_values/geneset_p_value.rds")

geneset_p_value_strong_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh/resample_ACAT_50_qc/p_values/geneset_p_value.rds")

geneset_p_value_maxstat <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh/resample_maxstat/p_values/geneset_p_value.rds")
p4<-gg_qqplot(rbind(geneset_p_value [grepl("NO_SITE",rownames(geneset_p_value)),]),title="GENESET: Power+ACAT (default qc)")
library(eulerr)


p5<-gg_qqplot(module_results_sceptre_NTC$acat_pval,title= "GENESET: Sceptre+ACAT (strong qc)")
p8<-gg_qqplot(module_results_sceptre_NTC_default$acat_pval,title= "GENESET: Sceptre+ACAT (default qc)")
p6<-gg_qqplot(rbind(geneset_p_value_maxstat [grepl("NO_SITE",rownames(geneset_p_value_maxstat)),]),title="GENESET: Power+Maxstat")


p7<-gg_qqplot(rbind(geneset_p_value_strong_qc [grepl("NO_SITE",rownames(geneset_p_value_strong_qc)),]),title="GENESET: Power+ACAT (strong qc)")
```





```{r,fig.width=18,fig.height=15}
grid.arrange(p1,p2,p3,ncol=1)
```

```{r,fig.width=10,fig.height=10}
grid.arrange(p4,p7,p8,p5,ncol=2)
```



```{r}
p4<-gg_qqplot_df(rbind(geneset_p_value [grepl("NO_SITE",rownames(geneset_p_value)),]),title="GENESET: Power+ACAT (default qc)")
library(eulerr)


p5<-gg_qqplot_df(module_results_sceptre_NTC$acat_pval,title= "GENESET: Sceptre+ACAT")

p6<-gg_qqplot_df(rbind(geneset_p_value_maxstat [grepl("NO_SITE",rownames(geneset_p_value_maxstat)),]),title="GENESET: Power+Maxstat")


p7<-gg_qqplot_df(rbind(geneset_p_value_strong_qc [grepl("NO_SITE",rownames(geneset_p_value_strong_qc)),]),title="GENESET: Power+ACAT (strong qc)")
```

```{r,fig.width=10,fig.height=10}
grid.arrange(p4,p7,ncol=1)
```

# Frangieh - control



<span style="color: red;"> It is worth noticing that different from other datasets, this dataset contains cell-cycle related information: s_score, g2m_score, and phase (according to the author, phase is calculated using s_score and g2m_score). In sceptre paper, the author take those additional covariate in their glm model, and use permutation to get p value.</span>

<span style="color: red;"> To imitate this, I also take this three additional covariate, and (use conditional resampling to obtain p values.) </span>

```{r}
cell_covariate <- readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/control/cell_covariate.rds")

head(cell_covariate )
```

```{r,echo=F,include=F}
grna_cutoff<-20

gene_module_index_matrix <-  readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/control/hallmark_set/gene_module_index_matrix.rds")

grna_target_data_frame_omit <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/control/grna_target_dataframe.rds"))

discovery_pairs_cis <-readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/control/discovery_pairs_cis.rds")

grna_assignment_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/control/hallmark_set/grna_assignment_matrix.rds")

max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh_control/resample_ACAT_7_qc/p_values/geneset_p_value.rds"))

max_stat <- max_stat[rowSums(grna_assignment_matrix)>grna_cutoff, ]


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/frangieh_control/hallmark_set_genes/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/frangieh/control/hallmark_set/perturbation_gene_7_qc.rds")
 


sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, colnames(sceptre_wide)%in%rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]
```





```{r,echo=F,include=F}
library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id == grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
  print(length(remove_genes))
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
  print("no genes removed")
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

max_stat <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh_control/resample_maxstat_7_qc/p_values/geneset_p_value.rds"))

max_stat <- max_stat[rowSums(grna_assignment_matrix)>grna_cutoff, ]


max_stat[max_stat==0]<- 1e-20
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")



```









```{r,include=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots

significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_levels]
)


significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_levels]
)
significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_levels]
)


  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT` = significant_ids_gene_wise_p,
    `Power + ACAT `=significant_ids_max_stat,
 `Power + Maxstat`=significant_ids_max_stat_2
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_levels)
)
p1 <- euler_plot


```


```{r,include=F}
sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/frangieh_control/hallmark_set_genes/results_run_calibration_check.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- character(0)
 


sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, colnames(sceptre_wide)%in%rownames(gene_module_index_matrix), drop = FALSE]
sceptre_wide<-sceptre_wide  [grepl("NO_SITE",rownames(sceptre_wide )),]

module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id == grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC_default <- do.call(rbind, module_results)
module_results_sceptre_NTC_default <- module_results_sceptre_NTC_default[!is.na(module_results_sceptre_NTC_default$acat_pval), ]
```

```{r,include=F}
geneset_p_value_ACAT <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh_control/resample_ACAT_7_qc/p_values/geneset_p_value.rds")

geneset_p_value_Maxstat <- readRDS("/project/xuanyao/jiaming/Getting_started/output/snakemake/frangieh_control/resample_maxstat_7_qc/p_values/geneset_p_value.rds")

row_name <- rownames(geneset_p_value_Maxstat)
rownames(geneset_p_value_ACAT)<-seq(1,nrow(geneset_p_value_ACAT))

rownames(geneset_p_value_Maxstat)<-seq(1,nrow(geneset_p_value_Maxstat))

p2<-gg_qqplot_df(rbind(geneset_p_value_ACAT [grepl("SITE",row_name),]),title="GENESET: Power+ACAT")

p3<-gg_qqplot_df(rbind(geneset_p_value_Maxstat [grepl("SITE",row_name),]),title="GENESET: Power+maxstat")

grna_groups <- split(module_results_sceptre_NTC_default$acat_pval, module_results_sceptre_NTC_default$grna_id)

names(grna_groups)<-1:length(grna_groups)
p4<-gg_qqplot_list(grna_groups,title="GENESET: Sceptre + ACAT")

p5<-gg_qqplot(rbind(geneset_p_value_ACAT [grepl("SITE",row_name),]),title="GENESET: Power+ACAT")
p6<-gg_qqplot(module_results_sceptre_NTC_default$acat_pval,title="GENESET: Sceptre+ACAT")
```

```{r,fig.width=18,fig.height=22}
grid.arrange(p1,p2,p3,p4,p5,p6,ncol=2)
```




