---
title: "Research log June"
author: "jliucx"
date: "2025-6-1"
output: 
  workflowr::wflow_html:
    number_sections: true
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72

---
```{r,include=F}

setwd("/project/xuanyao/jiaming/Getting_started")

library(ggplot2)
library(dplyr)
library(readr)
library(tibble)
library(tidyr)
library(presto)
library(sceptre)
library(sceptredata)
library(readr)
library(dplyr)
library(Matrix)
library(parallel)
library(BH)
library(Rcpp)
library(peakRAM)
library(sgt)
library(heavytailcombtest)
library(gridExtra)
library(eulerr)
library(ggplot2)
library(dplyr)
library(readr)
library(tibble)
library(ACAT)
library(MASS)
source("code/functions_optimized.R")


gg_qqplot <- function(ps, ci = 0.95, title = "QQ Plot") {
  n <- length(ps)
  df <- data.frame(
    observed = pmin(-log10(sort(ps)), 6), # Cap observed values at 4
    expected = -log10(ppoints(n)),
    clower = pmin(-log10(qbeta(p = (1 - ci) / 2, shape1 = 1:n, shape2 = n:1)), 6), # Cap clower values at 4
    cupper = pmin(-log10(qbeta(p = (1 + ci) / 2, shape1 = 1:n, shape2 = n:1)), 6)  # Cap cupper values at 4
  )

  log10Pe <- expression(paste("Expected -log"[10], plain(P)))
  log10Po <- expression(paste("Observed -log"[10], plain(P)))

  ggplot(df) +
    geom_ribbon(mapping = aes(x = expected, ymin = clower, ymax = cupper), alpha = 0.1) +
    geom_point(aes(expected, observed), size = 1.5, color = "red") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.5) +
    xlab(log10Pe) + 
    ylab(log10Po) +
    ylim(0, 6) + # Limit y-axis to 0-5
    ggtitle(title) # Add title
}
```



# Morris cNMF




## Strong perturbation-gene qc

```{r,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris/program_data_nature_cNMF/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_official_应该没啥问题了/resample_normal_test_statistics.csv",row.names = 1)


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
 perturbation_gene_strong_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc.rds")

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_strong_qc)) {
  remove_genes <- perturbation_gene_strong_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_strong_qc)) {
  remove_genes <- perturbation_gene_strong_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]
```




```{r,echo=F}

max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_with_bug/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov/cNMF_3000_skew_t_approximation/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-5
max_stat_long_3 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3 <- max_stat_long_3 %>% filter(!is.na(p_value))


max_stat_long_3 <- max_stat_long_3 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_3$bh_pval <-  p.adjust(max_stat_long_3$p_value, method = "BH")
max_stat_long_3$identifier <- paste(max_stat_long_3$grna_id, max_stat_long_3$module, sep = "_")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-5
max_stat_long_4 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_4 <- max_stat_long_4 %>% filter(!is.na(p_value))


max_stat_long_4 <- max_stat_long_4 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_4$bh_pval <-  p.adjust(max_stat_long_4$p_value, method = "BH")
max_stat_long_4$identifier <- paste(max_stat_long_4$grna_id, max_stat_long_4$module, sep = "_")
max_stat_long_4 <- max_stat_long_4[match(max_stat_long_2$identifier, max_stat_long_4$identifier), ]


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-5
max_stat_long_5 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_5 <- max_stat_long_5 %>% filter(!is.na(p_value))


max_stat_long_5 <- max_stat_long_5 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_5$bh_pval <-  p.adjust(max_stat_long_5$p_value, method = "BH")
max_stat_long_5$identifier <- paste(max_stat_long_5$grna_id, max_stat_long_5$module, sep = "_")
```


```{r,cache=F,fig.width=16,fig.height=16,echo=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)

significant_ids_max_stat_3 <- na.omit(
  max_stat_long_3$identifier[max_stat_long_3$bh_pval < significant_level]
)

significant_ids_max_stat_4 <- na.omit(
  max_stat_long_4$identifier[max_stat_long_4$bh_pval < significant_level]
)
significant_ids_max_stat_5 <- na.omit(
  max_stat_long_5$identifier[max_stat_long_5$bh_pval < significant_level]
)

  
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT(truncated)` = significant_ids_gene_wise_p,
 
        `10000 original` = significant_ids_max_stat_2,
    `10000 current filter`=significant_ids_max_stat,
        `3000 current filter normal`=significant_ids_max_stat_4
 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "blue", "yellow","green"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 1)

par(mfrow = c(1, 1))
```


## NTC

```{r,fig.width=18,fig.height=10,cache=F}
p1<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_official_应该没啥问题了/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 10000 no approximation")

p2<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 3000 normal approximation")

p3<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_with_bug/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="No filter, 10000 no approximation")

ntc_pvals <- module_results_sceptre_NTC$acat_pval
p4<-gg_qqplot(ntc_pvals, title="NTC-ACAT")
grid.arrange(p3,p1,p2,p4,ncol=2)
```








# Morris our NMF

## Strong perturbation-gene qc

```{r,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris/program_data_NMF_set_300_50/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/our_NMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
 perturbation_gene_strong_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc.rds")

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_strong_qc)) {
  remove_genes <- perturbation_gene_strong_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_strong_qc)) {
  remove_genes <- perturbation_gene_strong_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/our_NMF_3000_wilcox_normal_approx_sceptre_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")

```


```{r,cache=F,fig.width=10,fig.height=10,echo=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)



  
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT` = significant_ids_gene_wise_p,
 
        `3000 current filter normal` = significant_ids_max_stat_2,
    `10000 current filter`=significant_ids_max_stat
 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "blue", "yellow","green"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 1)

par(mfrow = c(1, 1))
```

## NTC

```{r,fig.width=21,fig.height=7}
p1<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/our_NMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 10000 no approximation")

p2<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/our_NMF_3000_wilcox_normal_approx_sceptre_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 3000 normal approximation")




ntc_pvals <- module_results_sceptre_NTC$acat_pval
p4<-gg_qqplot(ntc_pvals, title="NTC-ACAT")
grid.arrange(p1,p2,p4,ncol=3)
```

# NTC - Random gene set

```{r,fig.width=15,fig.height=7,echo=F}


p5 <- gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/NTC_complete_set/Strong_qc_FAKE_NON_OVERLAP_set_70genes_per_set_3000_normal_wilcox/resample_normal_test_statistics.csv",row.names = 1))), title="filter, normal apprximation, 100 Non-overlapping fake sets, 70 genes per set")

p6 <- gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/NTC_complete_set/Strong_qc_FAKE_NON_OVERLAP_set_70genes_per_set_10000_no_wilcox/resample_normal_test_statistics.csv",row.names = 1))), title="filter, no apprximation, 100 Non-overlapping fake sets, 70 genes per set")


p3<-  gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/NTC_complete_set/NONOVERLAP_FAKE_set_70genes_per_set_3000_normal_wilcox/resample_normal_test_statistics.csv",row.names = 1))), title="No filter, normal apprximation, 100 Non-overlapping fake sets, 70 genes per set")

p4<-  gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/NTC_complete_set/NONOVERLAP_FAKE_set_70genes_per_set_10000_wilcox/resample_normal_test_statistics.csv",row.names = 1))), title="No filter, noapprximation, 100 Non-overlapping fake sets, 70 genes per set")


grid.arrange(p5,p6,p3,p4,ncol=2)
```



# Morris - HALLMARK



## Strong perturbation - gene qc

```{r,echo=F,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/resample_normal_test_statistics.csv",row.names = 1)


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc_complete_gene.rds")
 
 genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/genewise_p_combined_all.rds"))

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov/HALLMARK_10000_no_approximation/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-4
max_stat_long_3 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3 <- max_stat_long_3 %>% filter(!is.na(p_value))


max_stat_long_3 <- max_stat_long_3 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_3$bh_pval <-  p.adjust(max_stat_long_3$p_value, method = "BH")
max_stat_long_3$identifier <- paste(max_stat_long_3$grna_id, max_stat_long_3$module, sep = "_")

sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]

module_results <- list()
genewise_p_combined_all[genewise_p_combined_all>=1]<- 1 -1e-4
genewise_p_combined_all[genewise_p_combined_all==0]<- 1e-4
# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t <- do.call(rbind, module_results)

module_results_t <- module_results_t[!is.na(module_results_t$acat_pval), ]

module_results_t_NTC_10000 <- module_results_t %>%
  filter(grepl("^NTC", grna_id))

module_results_t <- module_results_t %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

```





```{r,cache=F,echo=F}

joined_df <- inner_join(module_results_t, max_stat_long, by = c("grna_id", "module"))

p1000<-ggplot(joined_df, aes(x = -log10(acat_pval), y = -log10(p_value))) +
  geom_point(color = "black", alpha = 0.3, size = 1.5) +
  geom_abline(slope = 1, intercept = 0, color = "darkred", linetype = "solid", linewidth = 1) +
  labs(
    x = "wilcoxon+ACAT",
    y = "wilcoxon+ours",
    title = "Wilcoxon(10000)+ACAT vs Wilcoxon(10000)+Our"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )

```


```{r,cache=F,fig.width=8,fig.height=8,echo=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)

significant_ids_max_stat_3 <- na.omit(
  max_stat_long_3$identifier[max_stat_long_3$bh_pval < significant_level]
)
  
  # Prepare the sets
  venn_data <- list(
    `Sceptre+Cauchy` = significant_ids_gene_wise_p,
    `remove spike strong qc`=significant_ids_max_stat,
        `remove spike strong qc normal`=significant_ids_max_stat_3,
    `not remove`=significant_ids_max_stat_2
 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 1)

par(mfrow = c(1, 1))
```


## NTC

```{r,fig.width=15,fig.height=7,cache=F}


p1<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_official_应该没啥问题了/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 10000 no approximation")

p2<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 3000 normal approximation")

p3<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_with_bug/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="No filter, 10000 no approximation")



ntc_pvals <- module_results_sceptre_NTC$acat_pval
p4<-gg_qqplot(ntc_pvals, title="NTC-ACAT")
grid.arrange(p3,p1,p2,p4,ncol=2)

```



# June 7 : Wilcoxon ACAT vs Wilcoxon Our combination at different setting

## 3000 normal

```{r,echo=F,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/resample_normal_test_statistics.csv",row.names = 1)


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc_complete_gene.rds")
 
 genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_genewise_3000_wilcox_genewise_normal_approx_perturbation_gene_remove_strong_qc_rep/genewise_p_combined_all.rds"))
 
  genewise_p_combined_all[genewise_p_combined_all>=1]<- 1 - 1e-4
 

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")



sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]

module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t_3000 <- do.call(rbind, module_results)

module_results_t_3000 <- module_results_t_3000[!is.na(module_results_t_3000$acat_pval), ]

module_results_t_NTC_3000 <- module_results_t_3000 %>%
  filter(grepl("^NTC", grna_id))

module_results_t_3000 <- module_results_t_3000 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

```



```{r,cache=F,echo=F}

library(dplyr)
library(ggplot2)

# inner join
joined_df <- inner_join(module_results_t_3000, max_stat_long, by = c("grna_id", "module"))


joined_df <- joined_df %>%
  mutate(
    x = -log10(acat_pval),
    y = -log10(p_value),
    x_trunc = ifelse(x > 25, 25, x)
  )


p2 <- ggplot(joined_df, aes(x = x_trunc, y = y)) +
  geom_point(color = "black", alpha = 0.3, size = 1.5) +
  geom_abline(slope = 1, intercept = 0, color = "darkred", linetype = "solid", linewidth = 1) +
  labs(
    x = "wilcoxon+ACAT",
    y = "wilcoxon+ours",
    title = "Wilcoxon(3000 normal)+ACAT vs Wilcoxon(10000)+Our"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )


```



## 100000 resample


```{r,echo=F,include=F,eval=TRUE}

 
 genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/genewise_p_combined_all_complete.rds"))

  genewise_p_combined_all[genewise_p_combined_all>=1]<- 1-1e-5
 
module_results <- list()

for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t_100000 <- do.call(rbind, module_results)

module_results_t_100000 <- module_results_t_100000[!is.na(module_results_t_100000$acat_pval), ]

module_results_t_NTC_100000 <- module_results_t_100000 %>%
  filter(grepl("^NTC", grna_id))


module_results_t_100000 <- module_results_t_100000 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

```



```{r,cache=F,echo=F}
library(dplyr)
library(ggplot2)

# inner join
joined_df <- inner_join(module_results_t_100000, max_stat_long, by = c("grna_id", "module"))

# ggplot 画图
p3<- ggplot(joined_df, aes(x = -log10(acat_pval), y = -log10(p_value))) +
  geom_point(color = "black", alpha = 0.3, size = 1.5) +
  geom_abline(slope = 1, intercept = 0, color = "darkred", linetype = "solid", linewidth = 1) +
  labs(
    x = "wilcoxon+ACAT",
    y = "wilcoxon+ours",
    title = "Wilcoxon(100000)+ACAT vs Wilcoxon(10000)+Our"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )


```

## Summary 



```{r,echo=F}
pNTC_10000 <- gg_qqplot(module_results_t_NTC_10000$acat_pval, title="Wilcoxon(10000)+ACAT")
pNTC_3000 <- gg_qqplot(module_results_t_NTC_3000$acat_pval, title="Wilcoxon(3000 normal)+ACAT")
pNTC_100000 <- gg_qqplot(module_results_t_NTC_100000$acat_pval, title="Wilcoxon(100000)+ACAT")
```




```{r,echo=F,cache=F}
 genewise_p_combined_all_3000 <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_genewise_3000_wilcox_genewise_normal_approx_perturbation_gene_remove_strong_qc_rep/genewise_p_combined_all.rds"))
 
  genewise_p_combined_all_3000[genewise_p_combined_all_3000>=1]<- 1 - 1e-4
  
genewise_p_combined_all_ntc_3000 <- genewise_p_combined_all_3000[grepl("NTC", rownames(genewise_p_combined_all_3000)), ]

pNTC_gene_3000<-gg_qqplot(as.vector(as.matrix(genewise_p_combined_all_ntc_3000)),title="Wilcoxon(3000 normal) Genewise")



 genewise_p_combined_all_100000 <- as.data.frame(rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/genewise_p_combined_all.rds"),readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/genewise_p_combined_complement.rds")))

  genewise_p_combined_all_100000[genewise_p_combined_all_100000>=1]<- 1-1e-5
  
genewise_p_combined_all_ntc_100000 <- genewise_p_combined_all_100000[grepl("NTC", rownames(genewise_p_combined_all_100000)), ]

pNTC_gene_100000<-gg_qqplot(as.vector(as.matrix(genewise_p_combined_all_ntc_100000)),title="Wilcoxon(100000) Genewise")



 genewise_p_combined_all_10000 <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/genewise_p_combined_all.rds"))

  genewise_p_combined_all_10000[genewise_p_combined_all_10000>=1]<- 1-1e-5
  
genewise_p_combined_all_ntc_10000 <- genewise_p_combined_all_10000[grepl("NTC", rownames(genewise_p_combined_all_10000)), ]

pNTC_gene_10000<-gg_qqplot(as.vector(as.matrix(genewise_p_combined_all_ntc_10000)),title="Wilcoxon(10000) Genewise")

```

```{r,cache=F,fig.width=15,fig.height=15}
grid.arrange(p1000,p3,p2,pNTC_10000,pNTC_100000,pNTC_3000,pNTC_gene_10000,pNTC_gene_100000,pNTC_gene_3000, ncol = 3)

```





## 3000 normal vs 100000

```{r}

common_rows <- intersect(rownames(genewise_p_combined_all_3000),
                         rownames(genewise_p_combined_all_100000))

# 按相同顺序 subset
genewise_100000_sub <- genewise_p_combined_all_100000[common_rows, , drop = FALSE]
genewise_3000_sub  <- genewise_p_combined_all_3000[common_rows, , drop = FALSE]

library(ggplot2)

# 创建 data frame
plot_df <- data.frame(
  p_100000 = -log10(as.vector(as.matrix(genewise_100000_sub))),
  p_3000 = -log10(as.vector(as.matrix(genewise_3000_sub)))
)

# 画图
pdiscovery<-ggplot(plot_df, aes(x = p_100000, y = p_3000)) +
  geom_point(alpha = 0.3, size = 1.2) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "-log10 p (100000)",
    y = "-log10 p (3000 normal)",
    title = "Discovery Set Gene-wise comparison: 100000 vs 3000 normal "
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))



```


```{r}

common_rows <- intersect(rownames(genewise_p_combined_all_ntc_100000),
                         rownames(genewise_p_combined_all_ntc_3000))

# 按相同顺序 subset
ntc_100000_sub <- genewise_p_combined_all_ntc_100000[common_rows, , drop = FALSE]
ntc_3000_sub  <- genewise_p_combined_all_ntc_3000[common_rows, , drop = FALSE]

plot_df_ntc <- data.frame(
  p_100000 = -log10(as.vector(as.matrix(ntc_100000_sub))),
  p_3000 = -log10(as.vector(as.matrix(ntc_3000_sub)))
)

pNTC<-ggplot(plot_df_ntc, aes(x = p_100000, y = p_3000)) +
  geom_point(alpha = 0.3, size = 1.2) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "-log10 p (NTC 100000)",
    y = "-log10 p (NTC 3000)",
    title = "NTC gene-wise comparison: 100000 vs 3000 normal"
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r,cache=F,fig.width=15,fig.height=7}
grid.arrange(pdiscovery,pNTC, ncol = 2)

```


## 10000 vs 100000

```{r,cache=F,echo=F}
max_stat_10000 <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/resample_normal_test_statistics.csv",row.names = 1)

max_stat_100000 <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/resample_normal_test_statistics.csv",row.names = 1)

max_stat_3000 <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)

max_stat_3000_t <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_3000_wilcox_t_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)

# 提取两个 data frame 中以 "SNP" 开头的 rownames
snp_rows_10000 <- rownames(max_stat_10000)[startsWith(rownames(max_stat_10000), "SNP")]
snp_rows_100000 <- rownames(max_stat_100000)[startsWith(rownames(max_stat_100000), "SNP")]
snp_rows_3000 <- rownames(max_stat_3000)[startsWith(rownames(max_stat_3000), "SNP")]

# 取交集
common_rows <- intersect(snp_rows_10000, snp_rows_100000)

# 子集化两个数据集
max_stat_10000_sub <- max_stat_10000[common_rows, , drop = FALSE]
max_stat_100000_sub <- max_stat_100000[common_rows, , drop = FALSE]
max_stat_3000_sub <- max_stat_3000[common_rows, , drop = FALSE]


plot(-log10(as.vector(as.matrix(max_stat_10000_sub))),-log10(as.vector(as.matrix(max_stat_100000_sub))),xlab="Our 10000",ylab="Our 100000",main="10000 resample vs 100000 resample")

```



```{r,echo=F}
max_stat_long_10000 <- max_stat_10000%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

max_stat_long_10000 <- max_stat_long_10000 %>% filter(!is.na(p_value)) %>%
semi_join(module_results_sceptre, by = c("grna_id", "module"))


max_stat_long_100000 <- max_stat_100000%>%
  rownames_to_column(var = "grna_id") %>% 
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_100000 <- max_stat_long_100000 %>% filter(!is.na(p_value))%>%
semi_join(module_results_sceptre, by = c("grna_id", "module"))


max_stat_long_3000_t <- max_stat_3000_t%>%
  rownames_to_column(var = "grna_id") %>% 
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3000_t <- max_stat_long_3000_t %>% filter(!is.na(p_value))%>%
semi_join(module_results_sceptre, by = c("grna_id", "module"))

library(dplyr)
library(purrr)

# max_stat_long_100000 保留与 module_results_t_100000 匹配的行
max_stat_long_100000 <- max_stat_long_100000 %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

# module_results_t_3000 同样处理
module_results_t_3000 <- module_results_t_3000 %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

# module_results_sceptre 同样处理
module_results_sceptre <- module_results_sceptre %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

max_stat_long_10000 <- max_stat_long_10000 %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

max_stat_long_3000_t <- max_stat_long_3000_t %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

```

```{r,echo=F}
module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")


module_results_t_3000$bh_pval <- p.adjust(module_results_t_3000$acat_pval, method = "BH")
module_results_t_3000$identifier <- paste(module_results_t_3000$grna_id, module_results_t_3000$module, sep = "_")


module_results_t_100000$bh_pval <- p.adjust(module_results_t_100000$acat_pval, method = "BH")
module_results_t_100000$identifier <- paste(module_results_t_100000$grna_id, module_results_t_100000$module, sep = "_")


max_stat_long_100000$bh_pval <- p.adjust(max_stat_long_100000$p_value, method = "BH")
max_stat_long_100000$identifier <- paste(max_stat_long_100000$grna_id, max_stat_long_100000$module, sep = "_")


max_stat_long_10000$bh_pval <- p.adjust(max_stat_long_10000$p_value, method = "BH")
max_stat_long_10000$identifier <- paste(max_stat_long_10000$grna_id, max_stat_long_10000$module, sep = "_")

max_stat_long_3000_t$bh_pval <- p.adjust(max_stat_long_3000_t$p_value, method = "BH")
max_stat_long_3000_t$identifier <- paste(max_stat_long_3000_t$grna_id, max_stat_long_3000_t$module, sep = "_")

```





```{r,cache=F,fig.width=15,fig.height=15,echo=F}
library(eulerr)

significant_levels <- c(0.035,0.04,0.05,0.1)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method

significant_ids_max_stat_10000 <- na.omit(
  max_stat_long_10000$identifier[max_stat_long_10000$bh_pval < significant_level]
)

significant_ids_max_stat_100000 <- na.omit(
  max_stat_long_100000$identifier[max_stat_long_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_100000 <- na.omit(
  module_results_t_100000$identifier[module_results_t_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_3000 <- na.omit(
  module_results_t_3000$identifier[module_results_t_3000$bh_pval < significant_level]
)

significant_ids_sceptre_ACAT <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)


  
  # Prepare the sets
  venn_data <- list(

`Wilcoxon100000+ACAT(red)`=significant_ids_wilcoxon_ACAT_100000,

`Wilcoxon3000+ACAT(blue)`=significant_ids_wilcoxon_ACAT_3000,

`Sceptre+ACAT(yellow)`=significant_ids_sceptre_ACAT,

`Wilcoxon100000+Ours(green)`=significant_ids_max_stat_100000

 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "blue", "yellow","green"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 2)

```
```{r,cache=F,fig.width=15,fig.height=15,echo=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method

significant_ids_max_stat_10000 <- na.omit(
  max_stat_long_10000$identifier[max_stat_long_10000$bh_pval < significant_level]
)

significant_ids_max_stat_100000 <- na.omit(
  max_stat_long_100000$identifier[max_stat_long_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_100000 <- na.omit(
  module_results_t_100000$identifier[module_results_t_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_3000 <- na.omit(
  module_results_t_3000$identifier[module_results_t_3000$bh_pval < significant_level]
)

significant_ids_sceptre_ACAT <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_max_stat_3000_t <- na.omit(
  max_stat_long_3000_t$identifier[max_stat_long_3000_t$bh_pval < significant_level]
)
  
  # Prepare the sets
  venn_data <- list(
    
`Wilcoxon10000+Ours(yellow)`=significant_ids_max_stat_10000,

`Wilcoxon100000+Ours(green)`=significant_ids_max_stat_100000,

  `Wilcoxon3000+Ours+t (red)` = significant_ids_max_stat_3000_t

 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("yellow","green","red"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 1)

```

```{r,cache=F}
gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_3000_wilcox_t_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="Wilcoxon+Ours 3000 t approximation")
```

# June 23 : Wilcoxon ACAT vs Wilcoxon Our combination at different setting (without perturbation-gene qc)




```{r,echo=F,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")

perturbation_gene_qc<-character(0)  

discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- rbind(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/no_qc/HALLMARK_100000_wilcox_no_approx_perturbation_gene_no_perturbation_qc/resample_normal_test_statistics.csv",row.names = 1),read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/no_qc/HALLMARK_100000_wilcox_no_approx_perturbation_gene_no_perturbation_qc_p2/resample_normal_test_statistics.csv",row.names = 1))


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
 
genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_3000_wilcox_normal_approx_perturbation_gene_no_perturbation_qc_/genewise_p_combined_all.rds"))
 
genewise_p_combined_all[genewise_p_combined_all>=1]<- 1 - 1e-4
 

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]




library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")



module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t_3000 <- do.call(rbind, module_results)

module_results_t_3000 <- module_results_t_3000[!is.na(module_results_t_3000$acat_pval), ]

module_results_t_NTC_3000 <- module_results_t_3000 %>%
  filter(grepl("^NTC", grna_id))

module_results_t_3000 <- module_results_t_3000 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

```





```{r,echo=F,include=F,eval=TRUE}

 
 genewise_p_combined_all <- as.data.frame(rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_100000_wilcox_no_approx_perturbation_gene_no_perturbation_qc/genewise_p_combined_all.rds"),readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_100000_wilcox_no_approx_perturbation_gene_no_perturbation_qc_p2/genewise_p_combined_all.rds")))

  genewise_p_combined_all[genewise_p_combined_all>=1]<- 1-1e-5
 
module_results <- list()

for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t_100000 <- do.call(rbind, module_results)

module_results_t_100000 <- module_results_t_100000[!is.na(module_results_t_100000$acat_pval), ]

module_results_t_NTC_100000 <- module_results_t_100000 %>%
  filter(grepl("^NTC", grna_id))


module_results_t_100000 <- module_results_t_100000 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

```




```{r,echo=F}
max_stat_100000 <- rbind(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/no_qc/HALLMARK_100000_wilcox_no_approx_perturbation_gene_no_perturbation_qc/resample_normal_test_statistics.csv",row.names = 1),read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/no_qc/HALLMARK_100000_wilcox_no_approx_perturbation_gene_no_perturbation_qc_p2/resample_normal_test_statistics.csv",row.names = 1))


max_stat_long_100000 <- max_stat_100000%>%
  rownames_to_column(var = "grna_id") %>% 
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_100000 <- max_stat_long_100000 %>% filter(!is.na(p_value))%>%
semi_join(module_results_sceptre, by = c("grna_id", "module"))



library(dplyr)
library(purrr)

# max_stat_long_100000 保留与 module_results_t_100000 匹配的行
max_stat_long_100000 <- max_stat_long_100000 %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

# module_results_t_3000 同样处理
module_results_t_3000 <- module_results_t_3000 %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)



# module_results_sceptre 同样处理
module_results_sceptre <- module_results_sceptre %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)



```


```{r}
module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")


module_results_t_3000$bh_pval <- p.adjust(module_results_t_3000$acat_pval, method = "BH")
module_results_t_3000$identifier <- paste(module_results_t_3000$grna_id, module_results_t_3000$module, sep = "_")


module_results_t_100000$bh_pval <- p.adjust(module_results_t_100000$acat_pval, method = "BH")
module_results_t_100000$identifier <- paste(module_results_t_100000$grna_id, module_results_t_100000$module, sep = "_")


max_stat_long_100000$bh_pval <- p.adjust(max_stat_long_100000$p_value, method = "BH")
max_stat_long_100000$identifier <- paste(max_stat_long_100000$grna_id, max_stat_long_100000$module, sep = "_")



```





```{r,cache=F,fig.width=20,fig.height=20,echo=F}
library(eulerr)

significant_levels <- c(0.035,0.04,0.05,0.1)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method


significant_ids_max_stat_100000 <- na.omit(
  max_stat_long_100000$identifier[max_stat_long_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_100000 <- na.omit(
  module_results_t_100000$identifier[module_results_t_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_3000 <- na.omit(
  module_results_t_3000$identifier[module_results_t_3000$bh_pval < significant_level]
)

significant_ids_sceptre_ACAT <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)


  
  # Prepare the sets
  venn_data <- list(

`Wilcoxon100000+ACAT(red)`=significant_ids_wilcoxon_ACAT_100000,

`Wilcoxon3000+ACAT(blue)`=significant_ids_wilcoxon_ACAT_3000,

`Sceptre+ACAT(yellow)`=significant_ids_sceptre_ACAT,

`Wilcoxon100000+Ours(green)`=significant_ids_max_stat_100000

 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "blue", "yellow","green"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 2)

```

```{r,echo=F,cache=F}
max_stat_100000_no_qc <- rbind(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/no_qc/HALLMARK_100000_wilcox_no_approx_perturbation_gene_no_perturbation_qc/resample_normal_test_statistics.csv",row.names = 1),read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/no_qc/HALLMARK_100000_wilcox_no_approx_perturbation_gene_no_perturbation_qc_p2/resample_normal_test_statistics.csv",row.names = 1))

max_stat_100000_strong_qc <-  read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/resample_normal_test_statistics.csv",row.names = 1)

common <- intersect(rownames(max_stat_100000_no_qc), rownames(max_stat_100000_strong_qc))
no_qc2  <- max_stat_100000_no_qc[common, , drop = FALSE]
strong2 <- max_stat_100000_strong_qc[common, , drop = FALSE]

plot(as.vector(as.matrix(no_qc2)),as.vector(as.matrix(strong2)),main="100000 resample, with/without perturbation-gene qc",xlab="no genewise qc",ylab="with genewise qc")
```

 
# June 23 : Simulation

## Geneset 1

```{r,echo=F,include=F}
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)

base_dir <- "/project/xuanyao/jiaming/Getting_started/output/simulation/our_acat_comparison" 
percent_folders <- list.dirs(base_dir, recursive = FALSE)

# 筛选：100%、80%、50%开头的所有文件夹，或精确等于30%_with_effect
percent_folders <- percent_folders[
  grepl("/(100%|80%|50%|0%)", percent_folders) |
  grepl("/30%_with_effect", percent_folders)
]

# 初始化空数据框
all_data <- data.frame()

for (percent_folder in percent_folders) {
  percent_label <- basename(percent_folder)
  
  # 遍历子文件夹（例如 0.04, 0.045, ..., 0.07）
  subfolders <- list.dirs(percent_folder, recursive = FALSE)
  
  for (subfolder in subfolders) {
    effect_strength <- basename(subfolder)
    
    # 构造文件路径
    acat_path <- file.path(subfolder, "ACAT_p.csv")
    max_stat_path <- file.path(subfolder, "max_normalized_statistics.csv")
    
    if (file.exists(acat_path) && file.exists(max_stat_path)) {
      # 打印成功路径
      message("Reading: ", acat_path)
      message("         ", max_stat_path)

      # 读取CSV，取出p值
      acat_p <- read.csv(acat_path, row.names = 1)
      max_stat_p <- read.csv(max_stat_path, row.names = 1)
      
      # 统一行名
      common_rows <- intersect(rownames(acat_p), rownames(max_stat_p))
      acat_p <- acat_p[common_rows, , drop = FALSE]
      max_stat_p <- max_stat_p[common_rows, , drop = FALSE]
      
      # 整合成长格式
      df_long <- data.frame(
        id = common_rows,
        ACAT = acat_p[[1]],
        ours = max_stat_p[[1]]
      ) %>%
        pivot_longer(cols = c("ACAT", "ours"), names_to = "P_value_type", values_to = "p_value") %>%
        mutate(
          `-log10(p)` = -log10(p_value),
          Effect_Strength = effect_strength,
          Percentage = percent_label
        )
      
      all_data <- bind_rows(all_data, df_long)
    }
  }
}

```



```{r,echo=F}
# 生成新的 id，保证每两个 row 共用一个 id
new_ids <- paste0("guide_", rep(1:(nrow(all_data)/2), each = 2))

# 替换原有 id
all_data$id <- new_ids


summary_df <- all_data %>%
  mutate(
    Percentage_Short = str_extract(Percentage, "^\\d+%")
  ) %>%
  group_by(Percentage_Short, Effect_Strength) %>%
  summarise(n = n(), .groups = "drop")

summary_df



all_data <- all_data %>%
  mutate(Percentage_Short = str_extract(Percentage, "^\\d+%"))

# 拆分为 wide 格式以便配对
wide_data <- all_data %>%
  pivot_wider(
    id_cols = c(Percentage_Short, Effect_Strength, id),
    names_from = P_value_type,
    values_from = p_value
  )

# 确保 wide_data 中包含 both ACAT and ours 列
grouped_vectors <- wide_data %>%
  filter(!is.na(ACAT), !is.na(ours)) %>%
  group_by(Percentage_Short, Effect_Strength) %>%
  group_split() %>%
  setNames(map_chr(., ~ paste0(.x$Percentage_Short[1], "_", .x$Effect_Strength[1]))) %>%
  map(function(df) {
    sampled_df <- df
    list(
      ACAT = sampled_df$ACAT,
      ours = sampled_df$ours
    )
  })
lengths_summary <- map_df(grouped_vectors, function(lst) {
  tibble(
    ACAT_length = length(lst$ACAT),
    ours_length = length(lst$ours)
  )
}, .id = "Group")



```



```{r,cache=F,fig.width=12,fig.height=10,echo=F}
library(tidyverse)


summary_df <- map_dfr(names(grouped_vectors), function(group_name) {
  vec <- grouped_vectors[[group_name]]
  tibble(
    Group = group_name,
    Effect_Label = match(str_extract(group_name, "\\d\\.\\d{2,3}"), 
                         c("0.04", "0.045", "0.05", "0.055", "0.06", "0.065", "0.07")),
    Percentage = str_extract(group_name, "^\\d+%"),
    ACAT = sum(vec$ACAT < 0.05, na.rm = TRUE)/length(vec$ACAT),
    ours = sum(vec$ours < 0.05, na.rm = TRUE)/length(vec$ours)
  )
}) %>%
  pivot_longer(cols = c("ACAT", "ours"), names_to = "Method", values_to = "Count")
summary_df <- summary_df %>%
  mutate(Percentage = factor(Percentage, levels = c("0%","30%","50%", "80%", "100%"))) %>%
    mutate(
    Method = recode(Method, "ACAT" = "sceptre+ACAT")
  )

summary_df <- summary_df %>%
  mutate(
    Effect_Label = factor(Effect_Label, levels = 1:7)
  )
summary_df <- summary_df %>%
  mutate(
    Method = factor(Method, levels = c("sceptre+ACAT", "ours"))
  )
summary_df_0 <- summary_df %>%
  filter(Percentage == "0%") %>%
  group_by(Method) %>%
  summarise(
    Count = sum(Count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Group        = "0%_combined",
    # 把这两列换成和原表完全相同的 factor
    Effect_Label = factor(1,      levels = levels(summary_df$Effect_Label)),
    Percentage   = factor("0%",  levels = levels(summary_df$Percentage))
  )

# 2. 用 base R 重排 summary_df_0 的列顺序，确保跟原表一模一样
summary_df_0 <- summary_df_0[, c("Group", "Effect_Label", "Percentage", "Method", "Count")]

# 3. 把原表里非 0% 的行挑出来
summary_df_non0 <- summary_df[summary_df$Percentage != "0%", ]

# 4. 合并
summary_df_new <- rbind(summary_df_non0, summary_df_0)
ggplot(summary_df, aes(x = Effect_Label, y = Count, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  facet_wrap(~ Percentage, nrow = 1, scales = "free_x") +
  scale_x_discrete(labels = 1:7) +  
  labs(x = "Effect Strength", y = "power (significance 0.05)", fill = "P-value Type") +
  theme_minimal() +
  theme(
    panel.spacing.x = unit(1.5, "lines") 
  )



```

## Geneset 2

```{r,echo=F,include=F}
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)

base_dir <- "/project/xuanyao/jiaming/Getting_started/output/simulation/our_acat_comparison_p2" 
percent_folders <- list.dirs(base_dir, recursive = FALSE)

# 筛选：100%、80%、50%开头的所有文件夹，或精确等于30%_with_effect
percent_folders <- percent_folders[
  grepl("/(100%|80%|50%|0%)", percent_folders) |
  grepl("/30%_with_effect", percent_folders)
]

# 初始化空数据框
all_data <- data.frame()

for (percent_folder in percent_folders) {
  percent_label <- basename(percent_folder)
  
  # 遍历子文件夹（例如 0.04, 0.045, ..., 0.07）
  subfolders <- list.dirs(percent_folder, recursive = FALSE)
  
  for (subfolder in subfolders) {
    effect_strength <- basename(subfolder)
    
    # 构造文件路径
    acat_path <- file.path(subfolder, "ACAT_p.csv")
    max_stat_path <- file.path(subfolder, "max_normalized_statistics.csv")
    
    if (file.exists(acat_path) && file.exists(max_stat_path)) {
      # 打印成功路径
      message("Reading: ", acat_path)
      message("         ", max_stat_path)

      # 读取CSV，取出p值
      acat_p <- read.csv(acat_path, row.names = 1)
      max_stat_p <- read.csv(max_stat_path, row.names = 1)
      
      # 统一行名
      common_rows <- intersect(rownames(acat_p), rownames(max_stat_p))
      acat_p <- acat_p[common_rows, , drop = FALSE]
      max_stat_p <- max_stat_p[common_rows, , drop = FALSE]
      
      # 整合成长格式
      df_long <- data.frame(
        id = common_rows,
        ACAT = acat_p[[1]],
        ours = max_stat_p[[1]]
      ) %>%
        pivot_longer(cols = c("ACAT", "ours"), names_to = "P_value_type", values_to = "p_value") %>%
        mutate(
          `-log10(p)` = -log10(p_value),
          Effect_Strength = effect_strength,
          Percentage = percent_label
        )
      
      all_data <- bind_rows(all_data, df_long)
    }
  }
}

```




```{r,echo=F}
# 生成新的 id，保证每两个 row 共用一个 id
new_ids <- paste0("guide_", rep(1:(nrow(all_data)/2), each = 2))

# 替换原有 id
all_data$id <- new_ids


summary_df <- all_data %>%
  mutate(
    Percentage_Short = str_extract(Percentage, "^\\d+%")
  ) %>%
  group_by(Percentage_Short, Effect_Strength) %>%
  summarise(n = n(), .groups = "drop")


summary_df


all_data <- all_data %>%
  mutate(Percentage_Short = str_extract(Percentage, "^\\d+%"))


library(dplyr)

all_data <- all_data %>%
  mutate(
    Effect_Strength = if_else(Percentage_Short == "0%", 
                              "0.04", 
                              Effect_Strength)
  )





# 拆分为 wide 格式以便配对
wide_data <- all_data %>%
  pivot_wider(
    id_cols = c(Percentage_Short, Effect_Strength, id),
    names_from = P_value_type,
    values_from = p_value
  )

# 确保 wide_data 中包含 both ACAT and ours 列
grouped_vectors <- wide_data %>%
  filter(!is.na(ACAT), !is.na(ours)) %>%
  group_by(Percentage_Short, Effect_Strength) %>%
  group_split() %>%
  setNames(map_chr(., ~ paste0(.x$Percentage_Short[1], "_", .x$Effect_Strength[1]))) %>%
  map(function(df) {
    sampled_df <- df
    list(
      ACAT = sampled_df$ACAT,
      ours = sampled_df$ours
    )
  })
lengths_summary <- map_df(grouped_vectors, function(lst) {
  tibble(
    ACAT_length = length(lst$ACAT),
    ours_length = length(lst$ours)
  )
}, .id = "Group")


```



```{r,cache=F,fig.width=12,fig.height=10,echo=F}
library(tidyverse)


summary_df <- map_dfr(names(grouped_vectors), function(group_name) {
  vec <- grouped_vectors[[group_name]]
  tibble(
    Group = group_name,
    Effect_Label = match(str_extract(group_name, "\\d\\.\\d{2,3}"), 
                         c("0.04", "0.045", "0.05", "0.055", "0.06", "0.065", "0.07")),
    Percentage = str_extract(group_name, "^\\d+%"),
    ACAT = sum(vec$ACAT < 0.05, na.rm = TRUE)/length(vec$ACAT),
    ours = sum(vec$ours < 0.05, na.rm = TRUE)/length(vec$ours)
  )
}) %>%
  pivot_longer(cols = c("ACAT", "ours"), names_to = "Method", values_to = "Count")
summary_df <- summary_df %>%
  mutate(Percentage = factor(Percentage, levels = c("0%","30%","50%", "80%", "100%"))) %>%
    mutate(
    Method = recode(Method, "ACAT" = "sceptre+ACAT")
  )

summary_df <- summary_df %>%
  mutate(
    Effect_Label = factor(Effect_Label, levels = 1:7)
  )
summary_df <- summary_df %>%
  mutate(
    Method = factor(Method, levels = c("sceptre+ACAT", "ours"))
  )

ggplot(summary_df, aes(x = Effect_Label, y = Count, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  facet_wrap(~ Percentage, nrow = 1, scales = "free_x") +
  scale_x_discrete(labels = 1:7) +  
  labs(x = "Effect Strength", y = "power (significance 0.05)", fill = "P-value Type") +
  theme_minimal() +
  theme(
    panel.spacing.x = unit(1.5, "lines") 
  )



```




