---
title: "Research log June"
author: "jliucx"
date: "2025-6-1"
output: 
  workflowr::wflow_html:
    number_sections: true
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72

---
```{r,include=F}

setwd("/project/xuanyao/jiaming/Getting_started")

library(ggplot2)
library(dplyr)
library(readr)
library(tibble)
library(tidyr)
library(presto)
library(sceptre)
library(sceptredata)
library(readr)
library(dplyr)
library(Matrix)
library(parallel)
library(BH)
library(Rcpp)
library(peakRAM)
library(sgt)
library(heavytailcombtest)
library(gridExtra)
library(eulerr)
library(ggplot2)
library(dplyr)
library(readr)
library(tibble)
library(tidyr)
library(ACAT)
library(MASS)
source("code/functions_optimized.R")


gg_qqplot <- function(ps, ci = 0.95, title = "QQ Plot") {
  n <- length(ps)
  df <- data.frame(
    observed = pmin(-log10(sort(ps)), 6), # Cap observed values at 4
    expected = -log10(ppoints(n)),
    clower = pmin(-log10(qbeta(p = (1 - ci) / 2, shape1 = 1:n, shape2 = n:1)), 6), # Cap clower values at 4
    cupper = pmin(-log10(qbeta(p = (1 + ci) / 2, shape1 = 1:n, shape2 = n:1)), 6)  # Cap cupper values at 4
  )

  log10Pe <- expression(paste("Expected -log"[10], plain(P)))
  log10Po <- expression(paste("Observed -log"[10], plain(P)))

  ggplot(df) +
    geom_ribbon(mapping = aes(x = expected, ymin = clower, ymax = cupper), alpha = 0.1) +
    geom_point(aes(expected, observed), size = 1.5, color = "red") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.5) +
    xlab(log10Pe) + 
    ylab(log10Po) +
    ylim(0, 6) + # Limit y-axis to 0-5
    ggtitle(title) # Add title
}
```



# Morris cNMF




## Strong perturbation-gene qc

```{r,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris/program_data_nature_cNMF/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_official_应该没啥问题了/resample_normal_test_statistics.csv",row.names = 1)


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
 perturbation_gene_strong_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc.rds")

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_strong_qc)) {
  remove_genes <- perturbation_gene_strong_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_strong_qc)) {
  remove_genes <- perturbation_gene_strong_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]
```




```{r,echo=F}

max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_with_bug/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov/cNMF_3000_skew_t_approximation/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-5
max_stat_long_3 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3 <- max_stat_long_3 %>% filter(!is.na(p_value))


max_stat_long_3 <- max_stat_long_3 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_3$bh_pval <-  p.adjust(max_stat_long_3$p_value, method = "BH")
max_stat_long_3$identifier <- paste(max_stat_long_3$grna_id, max_stat_long_3$module, sep = "_")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-5
max_stat_long_4 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_4 <- max_stat_long_4 %>% filter(!is.na(p_value))


max_stat_long_4 <- max_stat_long_4 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_4$bh_pval <-  p.adjust(max_stat_long_4$p_value, method = "BH")
max_stat_long_4$identifier <- paste(max_stat_long_4$grna_id, max_stat_long_4$module, sep = "_")
max_stat_long_4 <- max_stat_long_4[match(max_stat_long_2$identifier, max_stat_long_4$identifier), ]


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-5
max_stat_long_5 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_5 <- max_stat_long_5 %>% filter(!is.na(p_value))


max_stat_long_5 <- max_stat_long_5 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_5$bh_pval <-  p.adjust(max_stat_long_5$p_value, method = "BH")
max_stat_long_5$identifier <- paste(max_stat_long_5$grna_id, max_stat_long_5$module, sep = "_")
```


```{r,cache=F,fig.width=16,fig.height=16,echo=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)

significant_ids_max_stat_3 <- na.omit(
  max_stat_long_3$identifier[max_stat_long_3$bh_pval < significant_level]
)

significant_ids_max_stat_4 <- na.omit(
  max_stat_long_4$identifier[max_stat_long_4$bh_pval < significant_level]
)
significant_ids_max_stat_5 <- na.omit(
  max_stat_long_5$identifier[max_stat_long_5$bh_pval < significant_level]
)

  
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT(truncated)` = significant_ids_gene_wise_p,
 
        `10000 original` = significant_ids_max_stat_2,
    `10000 current filter`=significant_ids_max_stat,
        `3000 current filter normal`=significant_ids_max_stat_4
 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "blue", "yellow","green"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 1)

par(mfrow = c(1, 1))
```


## NTC

```{r,fig.width=18,fig.height=10,cache=F}
p1<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_official_应该没啥问题了/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 10000 no approximation")

p2<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 3000 normal approximation")

p3<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_with_bug/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="No filter, 10000 no approximation")

ntc_pvals <- module_results_sceptre_NTC$acat_pval
p4<-gg_qqplot(ntc_pvals, title="NTC-ACAT")
grid.arrange(p3,p1,p2,p4,ncol=2)
```








# Morris our NMF

## Strong perturbation-gene qc

```{r,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris/program_data_NMF_set_300_50/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/our_NMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
 perturbation_gene_strong_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc.rds")

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_strong_qc)) {
  remove_genes <- perturbation_gene_strong_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_strong_qc)) {
  remove_genes <- perturbation_gene_strong_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/our_NMF_3000_wilcox_normal_approx_sceptre_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")

```


```{r,cache=F,fig.width=10,fig.height=10,echo=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)



  
  # Prepare the sets
  venn_data <- list(
    `Sceptre+ACAT` = significant_ids_gene_wise_p,
 
        `3000 current filter normal` = significant_ids_max_stat_2,
    `10000 current filter`=significant_ids_max_stat
 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "blue", "yellow","green"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 1)

par(mfrow = c(1, 1))
```

## NTC

```{r,fig.width=21,fig.height=7}
p1<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/our_NMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 10000 no approximation")

p2<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/our_NMF_3000_wilcox_normal_approx_sceptre_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 3000 normal approximation")




ntc_pvals <- module_results_sceptre_NTC$acat_pval
p4<-gg_qqplot(ntc_pvals, title="NTC-ACAT")
grid.arrange(p1,p2,p4,ncol=3)
```

# NTC - Random gene set

```{r,fig.width=15,fig.height=7}


p5 <- gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/NTC_complete_set/Strong_qc_FAKE_NON_OVERLAP_set_70genes_per_set_3000_normal_wilcox/resample_normal_test_statistics.csv",row.names = 1))), title="filter, normal apprximation, 100 Non-overlapping fake sets, 70 genes per set")

p6 <- gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/NTC_complete_set/Strong_qc_FAKE_NON_OVERLAP_set_70genes_per_set_10000_no_wilcox/resample_normal_test_statistics.csv",row.names = 1))), title="filter, no apprximation, 100 Non-overlapping fake sets, 70 genes per set")


p3<-  gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/NTC_complete_set/NONOVERLAP_FAKE_set_70genes_per_set_3000_normal_wilcox/resample_normal_test_statistics.csv",row.names = 1))), title="No filter, normal apprximation, 100 Non-overlapping fake sets, 70 genes per set")

p4<-  gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/NTC_complete_set/NONOVERLAP_FAKE_set_70genes_per_set_10000_wilcox/resample_normal_test_statistics.csv",row.names = 1))), title="No filter, noapprximation, 100 Non-overlapping fake sets, 70 genes per set")


grid.arrange(p5,p6,p3,p4,ncol=2)
```



# Morris - HALLMARK



## Strong perturbation - gene qc

```{r,echo=F,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/resample_normal_test_statistics.csv",row.names = 1)


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc_complete_gene.rds")
 
 genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/genewise_p_combined_all.rds"))

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value >=1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")

max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov/HALLMARK_10000_no_approximation/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-4
max_stat_long_2 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_2 <- max_stat_long_2 %>% filter(!is.na(p_value))


max_stat_long_2 <- max_stat_long_2 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_2$bh_pval <-  p.adjust(max_stat_long_2$p_value, method = "BH")
max_stat_long_2$identifier <- paste(max_stat_long_2$grna_id, max_stat_long_2$module, sep = "_")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)
max_stat[max_stat==0]<- 1e-4
max_stat_long_3 <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3 <- max_stat_long_3 %>% filter(!is.na(p_value))


max_stat_long_3 <- max_stat_long_3 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

max_stat_long_3$bh_pval <-  p.adjust(max_stat_long_3$p_value, method = "BH")
max_stat_long_3$identifier <- paste(max_stat_long_3$grna_id, max_stat_long_3$module, sep = "_")

sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]

module_results <- list()
genewise_p_combined_all[genewise_p_combined_all>=1]<- 1 -1e-4
genewise_p_combined_all[genewise_p_combined_all==0]<- 1e-4
# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t <- do.call(rbind, module_results)

module_results_t <- module_results_t[!is.na(module_results_t$acat_pval), ]

module_results_t_NTC_10000 <- module_results_t %>%
  filter(grepl("^NTC", grna_id))

module_results_t <- module_results_t %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

```





```{r,cache=F,echo=F}

joined_df <- inner_join(module_results_t, max_stat_long, by = c("grna_id", "module"))

p1000<-ggplot(joined_df, aes(x = -log10(acat_pval), y = -log10(p_value))) +
  geom_point(color = "black", alpha = 0.3, size = 1.5) +
  geom_abline(slope = 1, intercept = 0, color = "darkred", linetype = "solid", linewidth = 1) +
  labs(
    x = "wilcoxon+ACAT",
    y = "wilcoxon+ours",
    title = "Wilcoxon(10000)+ACAT vs Wilcoxon(10000)+Our"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )

```


```{r,cache=F,fig.width=8,fig.height=8,echo=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method
significant_ids_gene_wise_p <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_max_stat <- na.omit(
  max_stat_long$identifier[max_stat_long$bh_pval < significant_level]
)

significant_ids_max_stat_2 <- na.omit(
  max_stat_long_2$identifier[max_stat_long_2$bh_pval < significant_level]
)

significant_ids_max_stat_3 <- na.omit(
  max_stat_long_3$identifier[max_stat_long_3$bh_pval < significant_level]
)
  
  # Prepare the sets
  venn_data <- list(
    `Sceptre+Cauchy` = significant_ids_gene_wise_p,
    `remove spike strong qc`=significant_ids_max_stat,
        `remove spike strong qc normal`=significant_ids_max_stat_3,
    `not remove`=significant_ids_max_stat_2
 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "yellow", "orange","blue"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 1)

par(mfrow = c(1, 1))
```


## NTC

```{r,fig.width=15,fig.height=7,cache=F}


p1<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_strong_official_应该没啥问题了/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 10000 no approximation")

p2<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="current filter, 3000 normal approximation")

p3<-gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/cNMF_10000_wilcox_no_approx_sceptre_perturbation_gene_remove_with_bug/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="No filter, 10000 no approximation")



ntc_pvals <- module_results_sceptre_NTC$acat_pval
p4<-gg_qqplot(ntc_pvals, title="NTC-ACAT")
grid.arrange(p3,p1,p2,p4,ncol=2)

```



# June 7 : Wilcoxon ACAT vs Wilcoxon Our combination at different setting

## 3000 normal

```{r,echo=F,include=F}
gene_module_index_matrix <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/gene_module_index_matrix.rds")
grna_target_data_frame_omit <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/grna_target_data_frame_omit.rds")


discovery_pairs_cis <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data/discovery_pairs_cis.rds")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/resample_normal_test_statistics.csv",row.names = 1)


sceptre <- rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_discovery_analysis.rds")[,c(1,2,7)]
                 )
perturbation_gene_qc <- readRDS("/project/xuanyao/jiaming/Getting_started/data/Morris_adjusted/program_data_nature_cNMF/perturbation_gene_strong_qc_complete_gene.rds")
 
 genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_genewise_3000_wilcox_genewise_normal_approx_perturbation_gene_remove_strong_qc_rep/genewise_p_combined_all.rds"))
 
  genewise_p_combined_all[genewise_p_combined_all>=1]<- 1 - 1e-4
 

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
            module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre <- do.call(rbind, module_results)
module_results_sceptre <- module_results_sceptre[!is.na(module_results_sceptre$acat_pval), ]

library(tibble)

max_stat_long <- max_stat%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long <- max_stat_long %>% filter(!is.na(p_value))






max_stat_long <- max_stat_long %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))



max_stat_long$bh_pval <-  p.adjust(max_stat_long$p_value, method = "BH")
max_stat_long$identifier <- paste(max_stat_long$grna_id, max_stat_long$module, sep = "_")



sceptre <- readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/sceptre_all_pairs/author_threshold_updated/results_run_calibration_check.rds")[,c(1,2,7)]

sceptre <- sceptre[!is.na(sceptre$p_value), ]
sceptre$p_value[sceptre$p_value == 1] <- 1 - 1e-6

sceptre_wide <- sceptre %>%
  pivot_wider(names_from = response_id, values_from = p_value) %>%
  as.data.frame()

# Set the row names to be the grna_id and remove the grna_id column
rownames(sceptre_wide) <- sceptre_wide$grna_id
sceptre_wide <- sceptre_wide[, -1]

sceptre_wide <- sceptre_wide[rownames(max_stat)[rownames(max_stat) %in% rownames(sceptre_wide)], ]

sceptre_wide <- sceptre_wide[, rownames(gene_module_index_matrix), drop = FALSE]

# sceptre_wide <- sceptre_wide[rowSums(grna_assignment_matrix_author_threshold[rownames(sceptre_wide),])>100,]


library(eulerr)




module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(sceptre_wide))) {
  grna_id <- rownames(sceptre_wide)[i]
  p_values_row <- sceptre_wide[i,] 
  
  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  # 空字符向量，不会影响 setdiff
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 
      haha<-length(module_genes)
        module_genes <- setdiff(module_genes, remove_genes)

      
    
    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  # Remove NAs
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)
        
#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_sceptre_NTC <- do.call(rbind, module_results)
module_results_sceptre_NTC <- module_results_sceptre_NTC[!is.na(module_results_sceptre_NTC$acat_pval), ]

module_results <- list()

# Iterate over each row of t_gene_p_no_approximation to compute ACAT p-values for each module
for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t_3000 <- do.call(rbind, module_results)

module_results_t_3000 <- module_results_t_3000[!is.na(module_results_t_3000$acat_pval), ]

module_results_t_NTC_3000 <- module_results_t_3000 %>%
  filter(grepl("^NTC", grna_id))

module_results_t_3000 <- module_results_t_3000 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

```



```{r,cache=F,echo=F}

library(dplyr)
library(ggplot2)

# inner join
joined_df <- inner_join(module_results_t_3000, max_stat_long, by = c("grna_id", "module"))


joined_df <- joined_df %>%
  mutate(
    x = -log10(acat_pval),
    y = -log10(p_value),
    x_trunc = ifelse(x > 25, 25, x)
  )


p2 <- ggplot(joined_df, aes(x = x_trunc, y = y)) +
  geom_point(color = "black", alpha = 0.3, size = 1.5) +
  geom_abline(slope = 1, intercept = 0, color = "darkred", linetype = "solid", linewidth = 1) +
  labs(
    x = "wilcoxon+ACAT",
    y = "wilcoxon+ours",
    title = "Wilcoxon(3000 normal)+ACAT vs Wilcoxon(10000)+Our"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )


```



## 100000 resample


```{r,echo=F,include=F,eval=TRUE}

 
 genewise_p_combined_all <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/genewise_p_combined_all_complete.rds"))

  genewise_p_combined_all[genewise_p_combined_all>=1]<- 1-1e-5
 
module_results <- list()

for (i in seq_len(nrow(genewise_p_combined_all ))) {
  

  grna_id <- rownames(genewise_p_combined_all)[i]
  p_values_row <- genewise_p_combined_all[i,] 
  

  # Initialize an empty list to store ACAT p-values and module names for each module
  module_acat_pvals <- c()
  module_names <- c()
  matched_targets <- grna_target_data_frame_omit[grna_target_data_frame_omit$grna_id %in% grna_id, ]$grna_target
  cis_genes <- discovery_pairs_cis[discovery_pairs_cis$grna_target == matched_targets, ]$response_id  
  
if (grna_id %in% rownames(perturbation_gene_qc)) {
  remove_genes <- perturbation_gene_qc[grna_id, 2][[1]]
} else {
  remove_genes <- character(0)  
  
  
  
  
}
  
  
  # Iterate over each module in gene_modules to combine p-values using ACAT
  for (j in 1:ncol(gene_module_index_matrix)) {
    
    
    module_name <- colnames(gene_module_index_matrix)[j]  # Get the module name

    module_genes <- rownames(gene_module_index_matrix)[gene_module_index_matrix[, j] == 1]
module_genes <- intersect(module_genes, colnames(p_values_row))
    
      module_genes <- setdiff(module_genes, cis_genes) 

            module_genes <- setdiff(module_genes, remove_genes)


    # Only apply ACAT if there are genes to combine and they are not NA
    if (length(module_genes) > 0) {
      module_p_values <- as.numeric(p_values_row[module_genes])
      module_p_values <- module_p_values[!is.na(module_p_values)]  
      
      
      if (length(module_p_values) > 0) {
        acat_pval <- ACAT(module_p_values)

#      acat_pval <- combination.test(
#   module_p_values[!is.na(module_p_values)],
#   weights = NA,
#   method = "Cauchy",
#   tail.idx = 1
# )


        module_acat_pvals <- c(module_acat_pvals, acat_pval)
        module_names <- c(module_names, module_name)
      } else {
        # print("dont have valid p values for that module")
        module_acat_pvals <- c(module_acat_pvals, NA)
        module_names <- c(module_names, module_name)
      }
    } else {
      
      # print("dont have valid gene in that module")
      module_acat_pvals <- c(module_acat_pvals, NA)
      module_names <- c(module_names, module_name)
    }
  }
  
  # Store results as a data frame in the list
  module_results[[i]] <- data.frame(
    grna_id = grna_id,
    module = module_names,
    acat_pval = module_acat_pvals
  )
}


# Combine all results into a single data frame
module_results_t_100000 <- do.call(rbind, module_results)

module_results_t_100000 <- module_results_t_100000[!is.na(module_results_t_100000$acat_pval), ]

module_results_t_NTC_100000 <- module_results_t_100000 %>%
  filter(grepl("^NTC", grna_id))


module_results_t_100000 <- module_results_t_100000 %>%
  semi_join(module_results_sceptre, by = c("grna_id", "module"))

```



```{r,cache=F,echo=F}
library(dplyr)
library(ggplot2)

# inner join
joined_df <- inner_join(module_results_t_100000, max_stat_long, by = c("grna_id", "module"))

# ggplot 画图
p3<- ggplot(joined_df, aes(x = -log10(acat_pval), y = -log10(p_value))) +
  geom_point(color = "black", alpha = 0.3, size = 1.5) +
  geom_abline(slope = 1, intercept = 0, color = "darkred", linetype = "solid", linewidth = 1) +
  labs(
    x = "wilcoxon+ACAT",
    y = "wilcoxon+ours",
    title = "Wilcoxon(100000)+ACAT vs Wilcoxon(10000)+Our"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )


```

## Summary 



```{r,echo=F}
pNTC_10000 <- gg_qqplot(module_results_t_NTC_10000$acat_pval, title="Wilcoxon(10000)+ACAT")
pNTC_3000 <- gg_qqplot(module_results_t_NTC_3000$acat_pval, title="Wilcoxon(3000 normal)+ACAT")
pNTC_100000 <- gg_qqplot(module_results_t_NTC_100000$acat_pval, title="Wilcoxon(100000)+ACAT")
```




```{r,echo=F,cache=F}
 genewise_p_combined_all_3000 <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_genewise_3000_wilcox_genewise_normal_approx_perturbation_gene_remove_strong_qc_rep/genewise_p_combined_all.rds"))
 
  genewise_p_combined_all_3000[genewise_p_combined_all_3000>=1]<- 1 - 1e-4
  
genewise_p_combined_all_ntc_3000 <- genewise_p_combined_all_3000[grepl("NTC", rownames(genewise_p_combined_all_3000)), ]

pNTC_gene_3000<-gg_qqplot(as.vector(as.matrix(genewise_p_combined_all_ntc_3000)),title="Wilcoxon(3000 normal) Genewise")



 genewise_p_combined_all_100000 <- as.data.frame(rbind(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/genewise_p_combined_all.rds"),readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/genewise_p_combined_complement.rds")))

  genewise_p_combined_all_100000[genewise_p_combined_all_100000>=1]<- 1-1e-5
  
genewise_p_combined_all_ntc_100000 <- genewise_p_combined_all_100000[grepl("NTC", rownames(genewise_p_combined_all_100000)), ]

pNTC_gene_100000<-gg_qqplot(as.vector(as.matrix(genewise_p_combined_all_ntc_100000)),title="Wilcoxon(100000) Genewise")



 genewise_p_combined_all_10000 <- as.data.frame(readRDS("/project/xuanyao/jiaming/Getting_started/output/Morris/debug/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/genewise_p_combined_all.rds"))

  genewise_p_combined_all_10000[genewise_p_combined_all_10000>=1]<- 1-1e-5
  
genewise_p_combined_all_ntc_10000 <- genewise_p_combined_all_10000[grepl("NTC", rownames(genewise_p_combined_all_10000)), ]

pNTC_gene_10000<-gg_qqplot(as.vector(as.matrix(genewise_p_combined_all_ntc_10000)),title="Wilcoxon(10000) Genewise")

```

```{r,cache=F,fig.width=15,fig.height=15}
grid.arrange(p1000,p3,p2,pNTC_10000,pNTC_100000,pNTC_3000,pNTC_gene_10000,pNTC_gene_100000,pNTC_gene_3000, ncol = 3)

```





## 3000 normal vs 100000

```{r}

common_rows <- intersect(rownames(genewise_p_combined_all_3000),
                         rownames(genewise_p_combined_all_100000))

# 按相同顺序 subset
genewise_100000_sub <- genewise_p_combined_all_100000[common_rows, , drop = FALSE]
genewise_3000_sub  <- genewise_p_combined_all_3000[common_rows, , drop = FALSE]

library(ggplot2)

# 创建 data frame
plot_df <- data.frame(
  p_100000 = -log10(as.vector(as.matrix(genewise_100000_sub))),
  p_3000 = -log10(as.vector(as.matrix(genewise_3000_sub)))
)

# 画图
pdiscovery<-ggplot(plot_df, aes(x = p_100000, y = p_3000)) +
  geom_point(alpha = 0.3, size = 1.2) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "-log10 p (100000)",
    y = "-log10 p (3000 normal)",
    title = "Discovery Set Gene-wise comparison: 100000 vs 3000 normal "
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))



```


```{r}

common_rows <- intersect(rownames(genewise_p_combined_all_ntc_100000),
                         rownames(genewise_p_combined_all_ntc_3000))

# 按相同顺序 subset
ntc_100000_sub <- genewise_p_combined_all_ntc_100000[common_rows, , drop = FALSE]
ntc_3000_sub  <- genewise_p_combined_all_ntc_3000[common_rows, , drop = FALSE]

plot_df_ntc <- data.frame(
  p_100000 = -log10(as.vector(as.matrix(ntc_100000_sub))),
  p_3000 = -log10(as.vector(as.matrix(ntc_3000_sub)))
)

pNTC<-ggplot(plot_df_ntc, aes(x = p_100000, y = p_3000)) +
  geom_point(alpha = 0.3, size = 1.2) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "-log10 p (NTC 100000)",
    y = "-log10 p (NTC 3000)",
    title = "NTC gene-wise comparison: 100000 vs 3000 normal"
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r,cache=F,fig.width=15,fig.height=7}
grid.arrange(pdiscovery,pNTC, ncol = 2)

```


## 10000 vs 100000

```{r,cache=F,echo=F}
max_stat_10000 <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_10000_wilcox_no_approx_perturbation_gene_remove_strong_official/resample_normal_test_statistics.csv",row.names = 1)

max_stat_100000 <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_100000_wilcox_no_approx_perturbation_gene_remove_strong_qc_v/resample_normal_test_statistics.csv",row.names = 1)

max_stat_3000 <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_3000_wilcox_normal_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)

max_stat_3000_t <- read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_3000_wilcox_t_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)

# 提取两个 data frame 中以 "SNP" 开头的 rownames
snp_rows_10000 <- rownames(max_stat_10000)[startsWith(rownames(max_stat_10000), "SNP")]
snp_rows_100000 <- rownames(max_stat_100000)[startsWith(rownames(max_stat_100000), "SNP")]
snp_rows_3000 <- rownames(max_stat_3000)[startsWith(rownames(max_stat_3000), "SNP")]

# 取交集
common_rows <- intersect(snp_rows_10000, snp_rows_100000)

# 子集化两个数据集
max_stat_10000_sub <- max_stat_10000[common_rows, , drop = FALSE]
max_stat_100000_sub <- max_stat_100000[common_rows, , drop = FALSE]
max_stat_3000_sub <- max_stat_3000[common_rows, , drop = FALSE]


plot(-log10(as.vector(as.matrix(max_stat_10000_sub))),-log10(as.vector(as.matrix(max_stat_100000_sub))),xlab="Our 10000",ylab="Our 100000",main="10000 resample vs 100000 resample")

```



```{r,echo=F}
max_stat_long_10000 <- max_stat_10000%>%
  rownames_to_column(var = "grna_id") %>% # Add row names as a column
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

max_stat_long_10000 <- max_stat_long_10000 %>% filter(!is.na(p_value)) %>%
semi_join(module_results_sceptre, by = c("grna_id", "module"))


max_stat_long_100000 <- max_stat_100000%>%
  rownames_to_column(var = "grna_id") %>% 
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_100000 <- max_stat_long_100000 %>% filter(!is.na(p_value))%>%
semi_join(module_results_sceptre, by = c("grna_id", "module"))


max_stat_long_3000_t <- max_stat_3000_t%>%
  rownames_to_column(var = "grna_id") %>% 
  pivot_longer(cols = -grna_id, names_to = "module", values_to = "p_value")

# Remove rows with NA p-values
max_stat_long_3000_t <- max_stat_long_3000_t %>% filter(!is.na(p_value))%>%
semi_join(module_results_sceptre, by = c("grna_id", "module"))

library(dplyr)
library(purrr)

# max_stat_long_100000 保留与 module_results_t_100000 匹配的行
max_stat_long_100000 <- max_stat_long_100000 %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

# module_results_t_3000 同样处理
module_results_t_3000 <- module_results_t_3000 %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

# module_results_sceptre 同样处理
module_results_sceptre <- module_results_sceptre %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

max_stat_long_10000 <- max_stat_long_10000 %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

max_stat_long_3000_t <- max_stat_long_3000_t %>%
  semi_join(module_results_t_100000, by = c("grna_id", "module"))%>%
  arrange(grna_id, module)

```

```{r}
module_results_sceptre$bh_pval <- p.adjust(module_results_sceptre$acat_pval, method = "BH")
module_results_sceptre$identifier <- paste(module_results_sceptre$grna_id, module_results_sceptre$module, sep = "_")


module_results_t_3000$bh_pval <- p.adjust(module_results_t_3000$acat_pval, method = "BH")
module_results_t_3000$identifier <- paste(module_results_t_3000$grna_id, module_results_t_3000$module, sep = "_")


module_results_t_100000$bh_pval <- p.adjust(module_results_t_100000$acat_pval, method = "BH")
module_results_t_100000$identifier <- paste(module_results_t_100000$grna_id, module_results_t_100000$module, sep = "_")


max_stat_long_100000$bh_pval <- p.adjust(max_stat_long_100000$p_value, method = "BH")
max_stat_long_100000$identifier <- paste(max_stat_long_100000$grna_id, max_stat_long_100000$module, sep = "_")


max_stat_long_10000$bh_pval <- p.adjust(max_stat_long_10000$p_value, method = "BH")
max_stat_long_10000$identifier <- paste(max_stat_long_10000$grna_id, max_stat_long_10000$module, sep = "_")

max_stat_long_3000_t$bh_pval <- p.adjust(max_stat_long_3000_t$p_value, method = "BH")
max_stat_long_3000_t$identifier <- paste(max_stat_long_3000_t$grna_id, max_stat_long_3000_t$module, sep = "_")

```





```{r,cache=F,fig.width=15,fig.height=15,echo=F}
library(eulerr)

significant_levels <- c(0.035,0.04,0.05,0.1)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method

significant_ids_max_stat_10000 <- na.omit(
  max_stat_long_10000$identifier[max_stat_long_10000$bh_pval < significant_level]
)

significant_ids_max_stat_100000 <- na.omit(
  max_stat_long_100000$identifier[max_stat_long_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_100000 <- na.omit(
  module_results_t_100000$identifier[module_results_t_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_3000 <- na.omit(
  module_results_t_3000$identifier[module_results_t_3000$bh_pval < significant_level]
)

significant_ids_sceptre_ACAT <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)


  
  # Prepare the sets
  venn_data <- list(

`Wilcoxon100000+ACAT(red)`=significant_ids_wilcoxon_ACAT_100000,

`Wilcoxon3000+ACAT(blue)`=significant_ids_wilcoxon_ACAT_3000,

`Sceptre+ACAT(yellow)`=significant_ids_sceptre_ACAT,

`Wilcoxon100000+Ours(green)`=significant_ids_max_stat_100000

 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("red", "blue", "yellow","green"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 2)

```
```{r,cache=F,fig.width=15,fig.height=15,echo=F}
library(eulerr)

significant_levels <- c(0.05)
# significant_levels <- 0.01
plot_list <- list()

# Loop through significant levels and display plots
for (significant_level in significant_levels) {
  # Filter significant IDs for each method

significant_ids_max_stat_10000 <- na.omit(
  max_stat_long_10000$identifier[max_stat_long_10000$bh_pval < significant_level]
)

significant_ids_max_stat_100000 <- na.omit(
  max_stat_long_100000$identifier[max_stat_long_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_100000 <- na.omit(
  module_results_t_100000$identifier[module_results_t_100000$bh_pval < significant_level]
)

significant_ids_wilcoxon_ACAT_3000 <- na.omit(
  module_results_t_3000$identifier[module_results_t_3000$bh_pval < significant_level]
)

significant_ids_sceptre_ACAT <- na.omit(
  module_results_sceptre$identifier[module_results_sceptre$bh_pval < significant_level]
)

significant_ids_max_stat_3000_t <- na.omit(
  max_stat_long_3000_t$identifier[max_stat_long_3000_t$bh_pval < significant_level]
)
  
  # Prepare the sets
  venn_data <- list(
    
`Wilcoxon10000+Ours(yellow)`=significant_ids_max_stat_10000,

`Wilcoxon100000+Ours(green)`=significant_ids_max_stat_100000,

  `Wilcoxon3000+Ours+t (red)` = significant_ids_max_stat_3000_t

 
  )
  
  # Generate the Euler diagram and display it
euler_plot <- plot(
  euler(venn_data),
  fills = list(fill = c("yellow","green","red"), alpha = 0.5),
  edges = TRUE,
  quantities = TRUE,
  labels = list(font = 2, cex = 1.2, pos = 3), # 调整字体和位置
  main = paste("Significant level", significant_level)
)
plot_list[[length(plot_list) + 1]] <- euler_plot
}
grid.arrange(grobs = plot_list, ncol = 1)

```

```{r,cache=F}
gg_qqplot(as.vector(as.matrix(read.csv("/project/xuanyao/jiaming/Getting_started/output/Morris/remove_cis_match_grna/4_cov_high_expressed_gene/HALLMARK_3000_wilcox_t_approx_perturbation_gene_remove_strong_qc/resample_normal_test_statistics.csv",row.names = 1)[189:198,])), title="Wilcoxon+Ours 3000 t approximation")
```

# June 10 Simulation

## 100 random choosed gene as gene set

Random picked 100 qced genes from expression matrix of Morris data. And edit the expression of 700 cells. (I choose 700 because it is the mean number of affected cell per guide RNA in Morris data)

For the perturbed genes, their expressions is given by Binomial (Original, 1 - effect size).

I play with the number of perturbed genes (percentage of perturbed gene) and the effect size.

```{r,echo=F}
# 创建一个 2×6 的矩阵（这里初始化为 NA）
mat <- matrix(c(0.8735,0.812,0.02,9.9e-5,9.9e-5,9.9e-5,9.9e-5,0.8735,9.9e-5,9.9e-5,9.9e-5,9.9e-5,9.9e-5,9.9e-5), nrow = 2, ncol = 7,byrow=T)

# 指定行名
rownames(mat) <- c("1% perturbed gene", "5% perturbed gene")

# （可选）指定列名
colnames(mat) <- c(paste0("Effect 0.", 0:5) ,"Effect 0.8")

# 看一下结果
mat

```

```{r,echo=F,cache=F}

max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/simulation/Non-targeting_random_geneset/random_geneset_%1_size_0.2/resample_normal_test_statistics.csv",row.names = 1)

gg_qqplot(as.vector(as.matrix(max_stat)),title="Non-targeting (%1 genes effect size 0.2)")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/simulation/Non-targeting_random_geneset/random_geneset_%10_size_0.2/resample_normal_test_statistics.csv",row.names = 1)

gg_qqplot(as.vector(as.matrix(max_stat)),title="Non-targeting (%10 genes effect size 0.2)")


max_stat <- read.csv("/project/xuanyao/jiaming/Getting_started/output/simulation/Non-targeting_random_geneset/random_geneset_%10_size_0.5/resample_normal_test_statistics.csv",row.names = 1)

gg_qqplot(as.vector(as.matrix(max_stat)),title="Non-targeting (%10 genes effect size 0.5)")
```

